---
title: "Businesses"
format: html
execute:
  warning: false
---

# Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
library(swissparl)
library(furrr)
library(here)
```

# Download Businesses

For all businesses dealt with in Parliament between 2016 and 2025, businesses that were submitted before 2016 are also necessary. Therefore, I download all businesses from the first session in 2010 until the session of fall 2025.

```{r}
yr = 2010:2025
get_businesses <- function(yr) {
  folder <- here("Data/businesses")
  if (!dir.exists(folder)) {
    dir.create(folder)
  }

  out_file <- file.path(folder, paste0(yr, ".rds"))
  if (file.exists(out_file)) {
    return(NULL)
  } # skip existing files

  dt <- get_data(
    "Business",
    SubmissionDate = c(
      paste(">", yr, "-01-01", sep = ""),
      paste("<", yr, "-12-31", sep = "")
    ),
    Language = "EN"
  )

  write_rds(dt, file.path(folder, paste0(yr, ".rds")))
}

walk(
  yr,
  get_businesses,
  .progress = TRUE
)

businesses <- map_dfr(
  list.files(here("Data/businesses"), full.names = T),
  read_rds
)
write_rds(businesses, here("Data/businesses.rds"))
```

# Inspect Businesses
Now that I have all the businesses, I can start to inspect the data. Let's see what's inside the business tags.

```{r}
businesses |>
  separate_rows(Tags, TagNames, sep = "\\|") |>
  count(Tags, TagNames) |>
  print(n = Inf)
```

Take a close look inside one of the business tags (e.g. "Energie").

```{r}
businesses |>
  filter(str_detect(TagNames, "Energie")) |>
  count(Title) |>
  arrange(desc(n)) |>
  print(n = 10)
```

Insepct the different business statuses.

```{r}
businesses |>
  count(BusinessStatus, BusinessStatusText) |>
  arrange(n) |>
  print(n = Inf)
```

What about the business types?

```{r}
businesses |>
  count(BusinessType, BusinessTypeName) |>
  arrange(n) |>
  print(n = Inf)
```

# Clean Businesses

Here I do all the Cleaning of the business data frame and store it for later use. The filter I applied serves the purpose of reducing the amount of businesses and only keep those that can have debates in the parliament.

## Business Types

Here i filter **out** business types that have debates in the parliament. These are the following:

```{r}
businesses |> 
  filter(BusinessType %in% c(18, 19, 10, 8, 14)) |> 
  distinct(BusinessType, BusinessTypeName)
```

## Business Status

Here I filter **for** business statuses that with a high cchance already had a dabeate in the parliament. These are the following:

```{r}
businesses |> 
  filter(BusinessStatus %in% c(219, 222, 232, 229, 218, 215, 220)) |> 
  distinct(BusinessStatus, BusinessStatusText)
```

## Business Tags

There are businesses with tags that have a higher chance of being a business which is climate related. I flag them here so I can use this flag later when I will create a handcoding dataset. These are the following:

```{r}
businesses |> 
  separate_rows(Tags, TagNames, sep = "\\|") |>
  filter(Tags %in% c(52,48,66)) |> 
  distinct(Tags, TagNames)
```

## Business Details

There are other variables that hold business information that might be useful for later analysis. For this purpose I create variables that summarise this information. Next, I perform all the preprocessing steps and store the cleaned data frame.

```{r}
businesses_cleaned <- businesses |>
  filter(
    !(BusinessType %in% c(18, 19, 10, 8, 14)),
    BusinessStatus %in% c(219, 222, 232, 229, 218, 215, 220)
  ) |>
  mutate(
    business_details_short = ifelse(is.na(Description), SubmittedText, Description),
    business_details_long = paste(business_details_short, ReasonText),
    business_details_short_length = str_length(business_details_short),
    BusinessDetails_long_length = str_length(business_details_long),
    business_tag_climate = str_detect(Tags, "52|48|66"),
  )

write_rds(businesses_cleaned, here("Data/businesses_cleaned.rds"))
```