---
title: "Transcripts"
format: html
execute:
  warning: false
---

# Setup

Import required packages and helper functions.

```{r}
library(tidyverse)
library(swissparl)
library(furrr)
library(future)
# library(tidytext)
library(here)
source(here("Scripts/00_functions.R"))
```

# Import Data

```{r}
businesses_cleaned <- read_rds(here("Data/01_businesses_cleaned.rds"))
```

# Download Subjects

Download `SubjectBusiness` entries for each business, with automatic retries and exponential backoff on failure. Already downloaded files are skipped.

```{r}
get_subjects <- function(bsn, retries = 5, wait = 3) {
  folder <- here("Data/subjects")
  if (!dir.exists(folder)) {
    dir.create(folder)
  }

  out_file <- file.path(folder, paste0(bsn, ".rds"))
  if (file.exists(out_file)) {
    return(NULL)
  } # skip existing files

  for (attempt in 1:retries) {
    result <- try(
      get_data(
        table = "SubjectBusiness",
        BusinessShortNumber = bsn,
        Language = "DE"
      ),
      silent = TRUE
    )

    if (!inherits(result, "try-error")) {
      write_rds(result, out_file)
      # message("Downloaded ", bsn, " (attempt ", attempt, ")")
      return(TRUE)
    }

    # message("Error on ", bsn, " (attempt ", attempt, "): ", result)
    Sys.sleep(wait * attempt) # exponential backoff
  }

  # message("FAILED permanently: ", bsn)
  write(bsn, file = "failed_ids.txt", append = TRUE)
  return(FALSE)
}

plan(multisession, workers = 4)

# Start parallel jobs
fut <- future_walk(
  businesses_cleaned$BusinessShortNumber,
  get_subjects,
  .progress = TRUE
)

# WAIT UNTIL ALL JOBS ARE FINISHED
# value(fut)

plan(sequential)

if (!file.exists(here("Data/02_subjects.rds"))) {
  subjects <- map_dfr(
    list.files(here("Data/subjects"), full.names = TRUE),
    read_rds
  )
  write_rds(subjects, here("Data/02_subjects.rds"))
} else {
  subjects <- read_rds(here("Data/02_subjects.rds"))
}
```

Some businesses have no entries in `SubjectBusiness`, indicating they were never debated in either council.

```{r}
missing_bsn <- setdiff(
  businesses_cleaned$BusinessShortNumber,
  subjects$BusinessShortNumber
)
paste(length(missing_bsn), "businesses have no entries in SubjectBusiness.")
```

# Download Transcripts

Download transcript entries for each subject, again with retries and exponential backoff.

```{r}
get_transcripts <- function(sbj, retries = 5, wait = 3) {
  folder <- here("Data/transcripts")
  if (!dir.exists(folder)) {
    dir.create(folder)
  }

  out_file <- file.path(folder, paste0(sbj, ".rds"))
  if (file.exists(out_file)) {
    return(NULL)
  } # skip existing files

  for (attempt in 1:retries) {
    result <- try(
      get_data(
        table = "Transcript",
        IdSubject = as.double(sbj),
        Language = "DE"
      ),
      silent = TRUE
    )

    if (!inherits(result, "try-error")) {
      saveRDS(result, out_file)
      # message("Downloaded ", sbj, " (attempt ", attempt, ")")
      return(TRUE)
    }

    # message("Error on ", sbj, " (attempt ", attempt, "): ", result)
    Sys.sleep(wait * attempt) # exponential backoff
  }

  # message("FAILED permanently: ", sbj)
  write(sbj, file = "failed_ids.txt", append = TRUE)
  return(FALSE)
}

plan(multisession, workers = 4)

fut <- future_walk(
  subjects$IdSubject,
  get_transcripts,
  .progress = TRUE
)

# WAIT UNTIL ALL JOBS ARE FINISHED
# value(fut)

plan(sequential)

if (!file.exists(here("Data/02_transcripts.rds"))) {
  transcripts <- map_dfr(
    list.files(here("Data/transcripts"), full.names = TRUE),
    read_rds
  )
  write_rds(transcripts, here("Data/02_transcripts.rds"))
} else {
  transcripts <- read_rds(here("Data/02_transcripts.rds"))
}
```

# Inspect Transcripts

All meetings belong to either the National Council, the Council of States, or the United Federal Assembly. There are no committee pre-sessions. Good.

```{r}
transcripts |> count(MeetingCouncilAbbreviation)
```

Check speaker details — which councils do they come from? I want only parliamentarians, other speakers (especially from the federal council) will be filtered out. 

```{r}
transcripts |> count(CouncilId, CouncilName)
```

Since the analysis centers around contributions from parliamentarians that can be attributed to a specific party (i.e. Parliamentary Group), entries without assigned parliamentary groups will be filtered out. These typically include speakers from the federal council, but also a few negligible contributions from parliamentarians without assigned groups (e.g. independents or technical remarks).

```{r}
transcripts |> count(ParlGroupName, CouncilName)
```

Check the distribution of speaker functions. Entries where no parliamentarian is assigned are typically moderations or technical remarks and will be filtered out.

```{r}
transcripts |> count(SpeakerFunction) |> arrange(n)
```

For demonstration, here are some transcripts without assigned speaker functions.

```{r}
transcripts |> filter(is.na(SpeakerFunction)) |> pull(Text) |> sample(10)
```

Check the language distribution of the transcripts. Italian transcripts make only about 1% of the data and will be removed from the analysis for practical reasons.

```{r}
transcripts |> count(LanguageOfText) |> mutate(share = n / sum(n)) |> arrange(n)
```

Check vote entries — these will be filtered out since they hold no substantive content.

```{r}
transcripts |> filter(!is.na(VoteBusinessNumber)) |> pull(Text) |> sample(10)
```

Check entries with `[VS]` tags — these are presiding remarks not relevant to content analysis and will be removed. As follows are some examples for demonstration.

```{r}
transcripts |>
  filter(
    str_detect(Text, "\\[VS]")
  ) |> 
  pull(Text) |> 
  sample(10)
```

# Clean Transcripts

Filter transcripts to the relevant session range and remove entries that are not suitable for content analysis: Italian-language entries, Speakers outside National Council and Council of State (Especially Federal Council), vote records, unassigned speaker functions, and paragraphs with `[VS]` tags.

Clean the Data, so it is ready for further processing.

```{r}
sessions <- get_data(
  "Session",
  StartDate = c(">2015-01-01"),
  Language = "DE"
  )

sessions_cleaned <- sessions |>
  mutate(
    IdSession = as.character(ID), 
    legislative_period = as.character(LegislativePeriodNumber)
  ) |> 
  rename(session_start = StartDate, session_end = EndDate) |>
  select(IdSession, SessionName, TypeName, legislative_period, session_start, session_end)

transcripts_filtered <- transcripts |>
  left_join(sessions_cleaned, by = "IdSession") |> 
  filter(
    legislative_period >= 50, # start with legislative period 50, which started in late 2015
    IdSession <= 5210, # end with Autumn session 2025, which is the most recent session
    LanguageOfText != "IT", # exclude Italian-language entries
    CouncilId %in% c(1, 2), # keep only Speakers from the National Council and Council of States
    !is.na(ParlGroupName), # exclude entries without assigned parliamentary groups
    is.na(VoteBusinessNumber), # exclude vote records
    !is.na(SpeakerFunction), # exclude moderation and technical remarks
    !str_detect(Text, "\\[VS]") # exclude paragraphs with VS (Vorsitzende) tags
  )

businesses_climate <- businesses_cleaned |>
  filter(business_tag_climate == TRUE) |>
  distinct(BusinessShortNumber) |>
  pull(BusinessShortNumber)

subjects_climate <- subjects |>
  filter(BusinessShortNumber %in% businesses_climate) |>
  distinct(IdSubject) |>
  pull(IdSubject)

transcripts_cleaned <- transcripts_filtered |>
  rename(transcript_id = ID) |>
  mutate(
    transcript_text = Text |>
      str_replace_all("\\[PAGE \\d+\\]", "") |> # remove pagination
      str_replace_all("\\[GZ]", "") |> # remove [GZ] tags (major caesura)
      str_squish(),
    transcript_word_count = str_count(transcript_text),
    business_tag_climate = if_else(
      IdSubject %in% subjects_climate,
      TRUE,
      FALSE
    ),
    meeting_id = as.character(MeetingVerbalixOid),
    council_id = as.character(CouncilId),
    canton_id = as.character(CantonId),
    meeting_date = as.Date(Start),
    ParlGroupName = case_when(
      ParlGroupAbbreviation %in% c("M-E", "BD") ~ "Die Mitte-Fraktion / CVP-& BD-Fraktion",
      .default = ParlGroupName
    )
  ) |>
  select(
    !c(
      VoteId,
      VoteBusinessNumber,
      VoteBusinessShortNumber,
      VoteBusinessTitle,
      Type,
      SortOrder,
      Start,
      End,
      Modified,
      StartTimeWithTimezone,
      EndTimeWithTimezone
    )
  )

write_rds(transcripts_cleaned, here("Data/02_transcripts_cleaned.rds"))
```

# Check Cleaned Transcripts

Session statistics:

```{r}
# session_stats <-
transcripts_cleaned |>
  group_by(IdSession) |>
  summarise(
    number_of_meetings = n_distinct(meeting_id),
    number_of_transcripts = n_distinct(transcript_id),
    number_of_businesses = n_distinct(IdSubject)
  )
```

Meeting statistics:

```{r}
transcripts_cleaned |>
  group_by(MeetingVerbalixOid) |>
  summarise(
    number_of_transcripts = n_distinct(transcript_id),
    number_of_businesses = n_distinct(IdSubject)
  )
```

# Create Paragraphs

Extract individual paragraphs from transcript texts using HTML `<p>` tags, then clean inline HTML markup and assign unique paragraph IDs.

```{r}
paragraphs_raw <- transcripts_cleaned |>
  mutate(
    paragraph_text = str_extract_all(transcript_text, "(?<=<p>)(.*?)(?=</p>)")
  ) |>
  unnest(paragraph_text)

paragraphs_cleaned <- paragraphs_raw |>
  group_by(transcript_id) |>
  mutate(
    paragraph_text = str_replace_all(paragraph_text, "<[^>]+>", ""), # remove HTML tags for italics and bold etc.
    paragraph_id = paste0(transcript_id, "-", row_number())
  ) |>
  ungroup() |>
  mutate(
    paragraph_word_count = str_count(paragraph_text, "\\S+")
  )

write_rds(paragraphs_cleaned, here("Data/02_paragraphs_cleaned.rds"))
```

# Visualize

Plot paragraph length distribution:

```{r}
plot_word_count_distribution(
  data = paragraphs_cleaned,
  title = "Distribution of Paragraph Lengths in Complete Dataset",
  output_path = "Outputs/02_paragraphs_cleaned_text_length_distribution.png"
)
```

Plot paragraph language distribution:

```{r}
plot_categorical_distribution(
  data = paragraphs_cleaned,
  group_col = "LanguageOfText",
  x_label = "Language of Paragraphs",
  title = "Distribution of Languages in Complete Dataset",
  output_path = "Outputs/02_paragraphs_cleaned_language_distribution.png"
)
```

Plot business tag distribution:

```{r}
plot_categorical_distribution(
  data = paragraphs_cleaned,
  group_col = "business_tag_climate",
  x_label = "Energy, Transport or Environment Related Business",
  title = "Distribution of Business Tags in Complete Dataset",
  output_path = "Outputs/02_paragraphs_cleaned_business_tags_distribution.png"
)
```
