---
title: "Handcoding Data"
format: 
  html:
    embed-resources: true
execute:
  warning: false
---

# Setup

Import required packages and helper functions.

```{r}
library(tidyverse)
library(dotenv)
library(here)
source(here("Scripts/00_functions.R"))
load_dot_env(here(".env"))
```

# Import Data

```{r}
paragraphs_cleaned <- read_csv(here("Data/02_paragraphs_cleaned.csv"))
```

# Create Handcoding Dataset

Create a balanced sample of 500 paragraphs for manual labeling. The sampling targets:

- 50/50 distribution between French and German paragraphs
- 70/30 distribution between climate-related and non-climate businesses

To achieve this, paragraphs from climate-related businesses and French-language paragraphs are upweighted during sampling.

```{r}
set.seed(1234)
handcoding_dataset <- paragraphs_cleaned |>
  mutate(
    cb_weight = if_else(business_tag_climate == TRUE, 10, 1),
    lang_weight = if_else(LanguageOfText == "FR", 3, 1),
    weight = cb_weight * lang_weight
  ) |>
  slice_sample(
    n = 500,
    weight_by = weight
  )

write_rds(handcoding_dataset, here("Data/03_handcoding_dataset.rds"))
```

# Visualize

Plot paragraph length distribution.

```{r}
plot_word_count_distribution(
  data = handcoding_dataset,
  title = "Distribution of Paragraph Lengths in Handcoding Dataset",
  output_path = "Outputs/03_handcoding_dataset_text_length_distribution.png"
)
```

Plot language distribution.

```{r}
plot_categorical_distribution(
  data = handcoding_dataset,
  group_col = "LanguageOfText",
  x_label = "Language of Paragraphs",
  title = "Distribution of Languages in Handcoding Dataset",
  output_path = "Outputs/03_handcoding_dataset_language_distribution.png"
)
```

Plot business tag distribution.

```{r}
plot_categorical_distribution(
  data = handcoding_dataset,
  group_col = "business_tag_climate",
  x_label = "Energy, Transport or Environment Related Business",
  title = "Distribution of Business Tags in Handcoding Dataset",
  output_path = "Outputs/03_handcoding_dataset_business_tags_distribution.png"
)
```

# Translate Paragraphs

Translate all paragraphs (both French and German) to English using the DeepL API so that the handcoding dataset is in a single language. Credentials are loaded from the `.env` file.

```{r}
handcoding_dataset_full <- if (file.exists(here("Data/03_handcoding_translations.rds"))) {

  translations_df <- read_rds(here("Data/03_handcoding_translations.rds"))
  handcoding_dataset |> left_join(translations_df, by = join_by(paragraph_id))

} else {

  translations_vec <- map_chr(
    handcoding_dataset$paragraph_text,
    deepl_translate,
    .progress = TRUE
  )

  translations_df <- tibble(
    paragraph_id = handcoding_dataset$paragraph_id,
    paragraph_translated = translations_vec
  )

  write_rds(translations_df, here("Data/03_handcoding_translations.rds"))
  handcoding_dataset |> left_join(translations_df, by = join_by(paragraph_id))
}
```

# Export

Select relevant columns and export to CSV for labeling in Label Studio.

```{r}
handcoding_dataset_translated <- handcoding_dataset_full |>
  select(
    transcript_id,
    paragraph_id,
    business_tag_climate,
    LanguageOfText,
    paragraph_text,
    paragraph_translated
  )

write_csv(handcoding_dataset_translated, here("Data/03_handcoding_dataset_translated.csv"))
```

