---
title: "Dictionary Creation"
format: html
execute:
  warning: false
---

# Setup

Import required packages and helper functions.

```{r}
library(tidyverse)
library(dotenv)
library(ellmer)
library(furrr)
library(future)
# library(openxlsx)
library(here)
source(here("Scripts/00_functions.R"))
load_dot_env(here(".env"))
```

# Import Data

```{r}
labeled_dataset <- read_csv(here("labeled_dataset.csv"))
paragraphs_cleaned <- read_rds(here("Data/02_paragraphs_cleaned.rds"))
```

# Process Labeled Data

Join back the original paragraph text from `paragraphs_cleaned`, since the labeled dataset has only the translated paragraphs. Convert the sentiment labels to a binary `climate` variable and remove annotation metadata columns.

```{r}
labeled_dataset_cleaned <-
  inner_join(
    labeled_dataset,
    paragraphs_cleaned,
    by = join_by(paragraph_id)
  ) |>
  mutate(
    climate = case_when(
      sentiment == "non_climate" ~ FALSE,
      sentiment == "non_classifiable" ~ FALSE,
      sentiment == "climate" ~ TRUE
    )
  ) |>
  select(-sentiment)

write_csv(labeled_dataset_cleaned, here("Data/04_labeled_dataset_cleaned.csv"))
```

# Visualize

Plot paragraph length distribution.

```{r}
plot_word_count_distribution(
  data = labeled_dataset_cleaned,
  title = "Distribution of Paragraph Lengths in Labeled Paragraphs",
  output_path = "Outputs/04_labeled_paragraphs_text_length_distribution.png"
)
```

Plot language distribution.

```{r}
plot_categorical_distribution(
  data = labeled_dataset_cleaned,
  group_col = "LanguageOfText",
  title = "Distribution of Languages in Labeled Paragraphs",
  x_label = "Language of Paragraphs",
  output_path = "Outputs/04_labeled_paragraphs_language_distribution.png"
)
```

Plot business tag distribution.

```{r}
plot_categorical_distribution(
  data = labeled_dataset_cleaned,
  group_col = "business_tag_climate",
  x_label = "Energy, Transport or Environment Related Business",
  title = "Distribution of Business Tags in Labeled Paragraphs",
  output_path = "Outputs/04_labeled_paragraphs_business_tags_distribution.png"
)
```

Plot class label distribution.

```{r}
plot_categorical_distribution(
  data = labeled_dataset_cleaned,
  group_col = "climate",
  x_label = "Climate Paragraph",
  title = "Distribution of Climate-Related Paragraphs \n in Labeled Dataset",
  output_path = "Outputs/04_labeled_paragraphs_class_distribution.png"
)
```

# Keyword Extraction

Extract climate-related keywords from labeled paragraphs using an LLM. Only paragraphs labeled as climate-related are processed. The extraction prompt is loaded from a text file and each paragraph is processed in parallel.

```{r}
if (file.exists(here("Data/04_paragraphs_keywords.rds"))) {
  
  paragraphs_keywords <- read_rds(here("Data/04_paragraphs_keywords.rds"))
  
} else {
  
  labeled_data_climate <- labeled_dataset_cleaned |> filter(climate == TRUE)
  
  prompt <- paste(readLines(here("extraction_prompt.txt")), collapse = " ")
  
  process_text <- function(text, index) {
    chat <- chat_openai(
      model = "gpt-5-mini",
      system_prompt = prompt
    )
  
    if (is.na(text) || text == "") {
      # message("Skipping empty row ", index)
      return(NA_character_)
    }
  
    result <- tryCatch(
      {
        chat$chat(text)
      },
      error = function(e) {
        # message("Error processing row ", index, ": ", e$message)
        NA_character_
      }
    )
  
    return(result)
  }
  
  rows_to_process <- seq_len(nrow(labeled_data_climate))
  # cat("Total rows to process:", length(rows_to_process), "\n")
  # cat("Using parallel processing with", future::nbrOfWorkers(), "workers\n")

  # start parallel processing
  plan(multisession, workers = 4)

  keywords_vec <- future_map2_chr(
    labeled_data_climate$paragraph_text,
    rows_to_process,
    process_text,
    .progress = TRUE,
    .options = furrr_options(seed = TRUE)
  )

  # end parallel processing
  plan(sequential)
  
  paragraphs_keywords <- tibble(
    paragraph_id = labeled_data_climate$paragraph_id,
    paragraph_text = labeled_data_climate$paragraph_text,
    LanguageOfText = labeled_data_climate$LanguageOfText,
    keywords = keywords_vec
  )
  
  write_rds(paragraphs_keywords, here("Data/04_paragraphs_keywords.rds"))
}
```

# Dictionary Preparation

Compile the extracted keywords into separate German and French dictionaries, deduplicate and sort entries, then translate all keywords to English using the DeepL API. The resulting dictionaries are exported as Excel files for manual review and curation.

```{r}
# Compile German keywords
german_keywords <- paragraphs_keywords |>
  filter(LanguageOfText == "DE") |>
  pull(keywords) |>
  str_split(",") |>
  unlist() |>
  str_squish() |>
  sort() |>
  unique() #|>
  # discard(~ .x == "")

# Compile French keywords
french_keywords <- paragraphs_keywords |>
  filter(LanguageOfText == "FR") |>
  pull(keywords) |>
  str_split(",") |>
  unlist() |>
  str_squish() |>
  sort() |>
  unique() #|>
  # discard(~ .x == "")

# Translate with DeepL

german_dictionary <- tibble(
  keyword_german = german_keywords,
  keyword_english = NA
)

french_dictionary <- tibble(
  keyword_french = french_keywords,
  keyword_english = NA
)

german_dictionary$keyword_english <- map_chr(
  german_dictionary$keyword_german,
  deepl_translate,
  .progress = TRUE
)

french_dictionary$keyword_english <- map_chr(
  french_dictionary$keyword_french,
  deepl_translate,
  .progress = TRUE
)

# Export for manual curation
write.xlsx(german_dictionary, here("Data/04_german_dictionary.xlsx"))
write.xlsx(french_dictionary, here("Data/04_french_dictionary.xlsx"))

# After reviewing, save the cleaned versions as *_curated.csv
```
