---
title: "Text Classification"
format: html
execute:
  warning: false
---

# Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
# library(openxlsx)
library(tidytext)
library(stopwords)
library(here)
library(zoo)
library(dotenv)
library(ellmer)
library(furrr)
library(future)
source(here("Scripts/00_functions.R"))
load_dot_env(here(".env"))
```


# Import Data

```{r}
#| label: data-import

paragraphs_cleaned <- read_rds(here("Data/02_paragraphs_cleaned.rds"))
# businesses_cleaned <- 
#   read_rds(here("Data/01_businesses_cleaned.rds")) |> 
#   select(BusinessShortNumber, Title)
subjects <- 
  read_rds(here("Data/02_subjects.rds")) |> 
  select(IdSubject, BusinessShortNumber, Title)
french_dictionary_curated <- 
  read_csv(here("french_dictionary_curated.csv")) |>
  rename(keyword_original = keyword_french)
german_dictionary_curated <- 
  read_csv(here("german_dictionary_curated.csv")) |>
  rename(keyword_original = keyword_german)
```

# Data Preparation

## Dictionary

Combine the curated German and French dictionaries into a single lookup table. Keywords are lowercased, stripped of non-alphanumeric characters (except hyphens and apostrophes), and whitespace-normalised.

```{r}
#| label: dictionary

dictionary_all_cleaned <- german_dictionary_curated |>
  bind_rows(french_dictionary_curated) |>
  mutate(
    keyword_original = str_to_lower(keyword_original),
    keyword_original = str_replace_all(keyword_original, "[^[:alnum:]'-]", ""),
    keyword_original = str_squish(keyword_original)
  ) |> 
  distinct(keyword_original, keyword_english)

write_rds(dictionary_all_cleaned, here("Data/05_dictionary_all_cleaned.rds"))
```

## Text Pre-Processing

Tokenise the paragraph texts into individual words, apply the same normalisation as the dictionary, and remove German and French stopwords.

```{r}
#| label: preprocessing

words <- paragraphs_cleaned |>
  select(paragraph_id, paragraph_text) |>
  unnest_tokens(
    output = "word",
    input = "paragraph_text",
    token = "words",
    drop = FALSE
  )

german_stopwords <- c(
  stopwords::stopwords("de", source = "snowball"),
  "dass"
) |>
  sort()

french_stopwords <- stopwords::stopwords("fr", source = "snowball")

stopwords_all <- tibble(stopwords = c(german_stopwords, french_stopwords))

words_cleaned <- words |>
  mutate(
    word = str_to_lower(word),
    word = str_replace_all(word, "[^[:alnum:]'-]", ""),
    word = str_squish(word)
  ) |>
  anti_join(stopwords_all, by = join_by(word == stopwords))

write_rds(words_cleaned, here("Data/05_words_cleaned.rds"))
```

# Dictionary Classification

Match each word against the dictionary and aggregate results at the paragraph level to determine which paragraphs contain at least one keyword match.

```{r}
#| label: apply-dictionary

paragraphs_classified <- words_cleaned |>
  left_join(
    dictionary_all_cleaned, 
    by = join_by(word == keyword_original), 
    relationship = "many-to-one") |>
  group_by(paragraph_id, paragraph_text) |>
  summarise(
    keywords_found = paste(na.omit(unique(keyword_english)), collapse = ", "),
    n_keywords_found = sum(!is.na(keyword_english)),
    climate_paragraph = if_else(n_keywords_found > 0, TRUE, FALSE),
    .groups = "drop"
  ) |>
  left_join(paragraphs_cleaned, by = join_by(paragraph_id))

write_rds(paragraphs_classified, here("Data/05_paragraphs_classified.rds"))
```

Do some preprocessing to prepare for final classification.

```{r}
#| label: climate-businesses

# n_distinct(subjects$IdSubject)

paragraphs_subjects_long <- left_join(
  paragraphs_classified,
  subjects,
  by = join_by(IdSubject),
  relationship = "many-to-many"
) |> 
  select(-IdSubject)

# paragraphs_subjects_long |> select(-BusinessShortNumber, -Title) |> distinct()

paragraphs_subjects_cleaned <- paragraphs_subjects_long |> 
  group_by(across(c(-BusinessShortNumber, -Title))) |> 
  summarise(
    BusinessShortNumber = paste(BusinessShortNumber, collapse = " | "),
    Title = paste(Title, collapse = " | ")
  )

# paragraphs_subjects_businesses <- left_join(
#   paragraphs_subjects,
#   businesses_cleaned,
#   by = join_by(BusinessShortNumber),
#   relationship = "many-to-one"
# )

# climate_businesses <- 
#   # tibble(
#   # climate_business = 
#     paragraphs_subjects_long |>
#     filter(climate_paragraph == TRUE) |>
#     distinct(BusinessShortNumber) |>
#     pull(BusinessShortNumber)
#   # )

paragraphs_classified_wide <- paragraphs_subjects_cleaned |>
  # mutate(
  #   climate = map_lgl(str_split(BusinessShortNumber, " \\| "), ~ any(.x %in% climate_businesses))
  # ) |>
  distinct() |> 
  rename(paragraph_text = paragraph_text.x) |> 
  select(-paragraph_text.y)
```

Inspect the businesses according to their climate relevance.

A business is assigned a `climate_temp` label if at least one paragraph and 5% of its paragraphs match the dictionary or the paragraph references the climate dictionary explicitly. This threshold ensures that only businesses with substantive climate content are included.

```{r}
business_classification <- paragraphs_classified_wide |> 
  group_by(BusinessShortNumber, Title) |>
  summarise(
    n_paragraphs = n(),
    n_climate_paragraphs = sum(climate_paragraph),
    proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
    climate_temp = case_when(n_climate_paragraphs >= 1 & proportion_climate_paragraphs >= 0.05 ~ TRUE,
    .default = FALSE)
  )
```

In a next step, an LLM will be used to check the business Titles of climate_temp businesses and assign the final `climate` label. This is necessary to exclude false positives, i.e. businesses that contain some climate keywords but are not actually about climate topics.

```{r}
#| label: llm-classification

if (file.exists(here("Data/05_business_classification_final.rds"))) {
  
  business_classificatoin_final <- 
    read_rds(here("Data/05_business_classification_final.rds"))
  
} else {
  
  bn_clim <- business_classification |> filter(climate_temp == TRUE)
  
  prompt <- paste(readLines(here("label_review.txt")), collapse = " ")
  
  process_text <- function(text, index) {
    chat <- chat_openai(
      model = "gpt-5-mini",
      system_prompt = prompt
    )
  
    result <- tryCatch(
      {
        chat$chat(text)
      },
      error = function(e) {
        NA_character_
      }
    )
  
    return(result)
  }
  
  rows_to_process <- seq_len(nrow(bn_clim))

  # set up parallel processing for LLM calls
  plan(multisession, workers = 4)

  bn_clim_reviewed <- bn_clim |> 
    mutate(
      llm_review = future_map2_chr(
        Title,
        rows_to_process,
        process_text,
        .progress = TRUE,
        .options = furrr_options(seed = TRUE)
      )
    )

  # end parallel processing
  plan(sequential)
  
  business_classification_final <- business_classification |> 
    left_join(
      bn_clim_reviewed |> select(BusinessShortNumber, llm_review),
      by = join_by(BusinessShortNumber)
    ) |> 
    mutate(
      final_climate = case_when(
        climate_temp == TRUE & llm_review == "True Positive" ~ TRUE,
        climate_temp == TRUE & llm_review == "False Positive" ~ FALSE,
        .default = FALSE
      )
    ) |> 
    select(-climate_temp, -llm_review)
  
  write_rds(business_classification_final, here("Data/05_business_classification_final.rds"))
}
```

```{r}
#| label: final-climate

# paragraphs_classified_wide_final <- paragraphs_classified_wide |>
#   group_by(BusinessShortNumber) |>
#   mutate(
#     n_paragraphs = n(),
#     n_climate_paragraphs = sum(climate_paragraph),
#     proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
#     final_climate = case_when(
#       # (proportion_climate_paragraphs >= 0.1 & n_climate_paragraphs > 1) | climate_paragraph == TRUE ~ TRUE,
#       proportion_climate_paragraphs >= 0. & n_climate_paragraphs > 1 ~ TRUE,
#       .default = FALSE
#     )
#   ) |>
#   ungroup()

paragraphs_classified_wide_final <- paragraphs_classified_wide |>
  left_join(
    business_classification_final |> select(BusinessShortNumber, final_climate),
    by = join_by(BusinessShortNumber)
  )

write_rds(
  paragraphs_classified_wide_final,
  here("Data/05_paragraphs_classified_wide_final.rds"))
```

The climate label is therefore assigned at the business level: if a parliamentary business is considered climate-related, all debates around it are treated as climate-related.

## Climate Business Analysis

Print the largest climate relevant businesses.

```{r}
paragraphs_classified_wide_final |> 
  filter(final_climate == TRUE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(n_paragraphs)) |> 
  slice_head(n = 20) |> 
  knitr::kable()
```

Print the climate businesses around the median.

```{r}
paragraphs_classified_wide_final |> 
  filter(final_climate == TRUE) |> 
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |> 
  slice_min(abs(n_paragraphs - median(n_paragraphs)), n = 10) |> 
  knitr::kable()
```

Print the smallest climate businesses.

```{r}
paragraphs_classified_wide_final |> 
  filter(final_climate == TRUE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(n_paragraphs)) |>  
  slice_tail(n = 20) |> 
  knitr::kable()
```

Print the climate businesses just above the threshold of 10%

```{r}
paragraphs_classified_wide_final |> 
  filter(final_climate == TRUE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(proportion_climate_paragraphs) |>  
  slice_head(n = 20) |> 
  knitr::kable()
```

Print non-climate businesses just below the threshold of 10%

```{r}
paragraphs_classified_wide_final |> 
  filter(final_climate == FALSE, proportion_climate_paragraphs < 0.1) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(proportion_climate_paragraphs)) |>  
  slice_head(n = 20) |> 
  knitr::kable()
```

# Prepare Data for Topic Modeling

Prepare climate-related paragraphs and transcripts for topic modeling. Filter out very short texts to ensure meaningful topic extraction.

```{r}
transcripts_climate_summary <- 
  paragraphs_classified_wide_final |>
  filter(final_climate == TRUE) |> 
  group_by(transcript_id) |>
  summarise(
    transcript_text = paste(paragraph_text, collapse = " "),
    transcript_word_length = str_count(transcript_text, "\\S+"),
    .groups = "drop"
  ) |>
  filter(transcript_word_length >= 50)

write_csv(
  transcripts_climate_summary,
  here("Data/05_transcripts_climate_for_topic_modeling.csv")
)
```