---
title: "Text Classification"
format: html
execute:
  warning: false
---

# Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
library(openxlsx)
library(tidytext)
library(stopwords)
library(here)
library(zoo)
source(here("Scripts/00_functions.R"))
```

# Import Data

```{r}
#| label: data-import

paragraphs_cleaned <- read_rds(here("Data/02_paragraphs_cleaned.rds"))
# businesses_cleaned <- 
#   read_rds(here("Data/01_businesses_cleaned.rds")) |> 
#   select(BusinessShortNumber, Title)
subjects <- 
  read_rds(here("Data/02_subjects.rds")) |> 
  select(IdSubject, BusinessShortNumber, Title)
french_dictionary_curated <- 
  read.xlsx(here("french_dictionary_curated.xlsx")) |>
  rename(keyword_original = keyword_french)
german_dictionary_curated <- 
  read.xlsx(here("german_dictionary_curated.xlsx")) |>
  rename(keyword_original = keyword_german)
```

# Data Preparation

## Dictionary

Combine the curated German and French dictionaries into a single lookup table. Keywords are lowercased, stripped of non-alphanumeric characters (except hyphens and apostrophes), and whitespace-normalised.

```{r}
#| label: dictionary

dictionary_all_cleaned <- german_dictionary_curated |>
  bind_rows(french_dictionary_curated) |>
  mutate(
    keyword_original = str_to_lower(keyword_original),
    keyword_original = str_replace_all(keyword_original, "[^[:alnum:]'-]", ""),
    keyword_original = str_squish(keyword_original)
  )

write_rds(dictionary_all_cleaned, here("Data/05_dictionary_all_cleaned.rds"))
```

## Text Pre-Processing

Tokenise the paragraph texts into individual words, apply the same normalisation as the dictionary, and remove German and French stopwords.

```{r}
#| label: preprocessing

paragraphs_words <- paragraphs_cleaned |>
  select(paragraph_id, paragraph_text) |>
  unnest_tokens(
    output = "word",
    input = "paragraph_text",
    token = "words",
    drop = FALSE
  )

german_stopwords <- c(
  stopwords::stopwords("de", source = "snowball"),
  "dass"
) |>
  sort()

french_stopwords <- stopwords::stopwords("fr", source = "snowball")

stopwords_all <- tibble(stopwords = c(german_stopwords, french_stopwords))

paragraphs_words_cleaned <- paragraphs_words |>
  mutate(
    word = str_to_lower(word),
    word = str_replace_all(word, "[^[:alnum:]'-]", ""),
    word = str_squish(word)
  ) |>
  anti_join(stopwords_all, by = join_by(word == stopwords))

write_rds(paragraphs_words_cleaned, here("Data/05_paragraphs_words_cleaned.rds"))
```

# Dictionary Classification

Match each word against the dictionary and aggregate results at the paragraph level to determine which paragraphs contain at least one keyword match.

```{r}
#| label: apply-dictionary

paragraphs_classified <- paragraphs_words_cleaned |>
  left_join(dictionary_all_cleaned, by = join_by(word == keyword_original)) |>
  group_by(paragraph_id, paragraph_text) |>
  summarise(
    keywords_found = paste(na.omit(unique(keyword_english)), collapse = ", "),
    n_keywords_found = sum(!is.na(keyword_english)),
    climate_paragraph = if_else(n_keywords_found > 0, TRUE, FALSE),
    .groups = "drop"
  ) |>
  left_join(paragraphs_cleaned, by = join_by(paragraph_id))

write_rds(paragraphs_classified, here("Data/05_paragraphs_classified.rds"))
```

The final climate label is assigned in three steps. First, businesses that contain at least one climate paragraph are identified as climate businesses.

```{r}
#| label: climate-businesses

paragraphs_subjects <- left_join(
  paragraphs_classified,
  subjects,
  by = join_by(IdSubject),
  relationship = "many-to-many"
)

# paragraphs_subjects_businesses <- left_join(
#   paragraphs_subjects,
#   businesses_cleaned,
#   by = join_by(BusinessShortNumber),
#   relationship = "many-to-one"
# )

climate_businesses <- paragraphs_subjects |>
  filter(climate_paragraph == TRUE) |>
  distinct(BusinessShortNumber) |>
  pull(BusinessShortNumber)
```

Second, all paragraphs belonging to a climate business are marked as climate-relevant.

```{r}
#| label: climate-paragraphs

paragraphs_classified_wide <- paragraphs_subjects |>
  mutate(
    climate = case_when(
      BusinessShortNumber %in% climate_businesses ~ TRUE,
      .default = FALSE
    )
  ) |>
  rename(paragraph_text = paragraph_text.x)
```

Third, a business is assigned a `final_climate` label if at least 10% of its paragraphs match the dictionary. This threshold ensures that only businesses with substantive climate content are included.

```{r}
#| label: final-climate

paragraphs_classified_wide_final <- paragraphs_classified_wide |>
  group_by(BusinessShortNumber) |>
  mutate(
    n_paragraphs = n(),
    n_climate_paragraphs = sum(climate_paragraph),
    proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
    final_climate = case_when(
      proportion_climate_paragraphs >= 0.1 ~ TRUE,
      .default = FALSE
    )
  ) |>
  ungroup()

write_rds(
  paragraphs_classified_wide_final,
  here("Data/05_paragraphs_classified_wide_final.rds"))
```

The climate label is therefore assigned at the business level: if a parliamentary business is considered climate-related, all debates around it are treated as climate-related.

# Visualize

## Visualization Setup

Define party display order and colour palette for all subsequent plots.

```{r}
#| label: viz-setup

party_names_order <- c(
  "Sozialdemokratische Fraktion",
  "Grüne Fraktion",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion",
  "Grünliberale Fraktion",
  "FDP-Liberale Fraktion",
  "Fraktion der Schweizerischen Volkspartei"
)

party_colors <- c(
  "Sozialdemokratische Fraktion" = "#E4032E",
  "Grüne Fraktion" = "#84B547",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion" = "#FF8C00",
  "Grünliberale Fraktion" = "#650959ff",
  "FDP-Liberale Fraktion" = "#0571C1",
  "Fraktion der Schweizerischen Volkspartei" = "#00843D"
)
```

## Across Categories

Share of climate-relevant vs. non-relevant paragraphs across the full dataset.

```{r}
p <- plot_categorical_distribution(
  paragraphs_classified_wide_final,
  group_col = "final_climate",
  x_label = "Climate Relevant Paragraph",
  title = "Distribution of Climate Relevant Paragraphs",
  output_path = "Outputs/05_paragraphs_climate_distribution.png"
)
write_rds(p, here("Outputs/05_paragraphs_climate_distribution.rds"))
p
```

Share of climate-relevant vs. non-relevant businesses.

```{r}
df_grouped <- paragraphs_classified_wide_final |>
  distinct(BusinessShortNumber, final_climate)

p <- plot_categorical_distribution(
  df_grouped,
  group_col = "final_climate",
  x_label = "Climate Relevant Businesses",
  y_label = "Percentage of Businesses",
  label_suffix = "businesses",
  title = "Distribution of Climate Relevant Businesses",
  output_path = "Outputs/05_businesses_climate_distribution.png"
)
write_rds(p, here("Outputs/05_businesses_climate_distribution.rds"))
p
```

## Categories Over Time

In absolute terms, the number of climate-relevant vs. non-climate paragraphs over time.

```{r}
df_grouped <- paragraphs_classified_wide_final |>
  mutate(meeting_date = as.Date(Start)) |>
  group_by(meeting_date, final_climate) |>
  summarise(count = n(), .groups = "drop")

p <- ggplot(df_grouped, aes(x = meeting_date, y = count, color = final_climate)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Climate vs. Non-Climate Paragraphs over Time",
    x = "Date",
    y = "Number of Paragraphs",
    color = "Climate"
  ) +
  theme_minimal()
ggsave(here("Outputs/05_plot_timeseries_absolute.png"))
write_rds(p, here("Outputs/05_plot_timeseries_absolute.rds"))
p
```

In relative terms, the share of climate-relevant paragraphs over the last 10 years, with a 30-session rolling average.

```{r}
share <- paragraphs_classified_wide_final |>
  group_by(meeting_date) |>
  summarise(
    share_climate = sum(paragraph_word_count[final_climate == TRUE]) / sum(paragraph_word_count) * 100
  ) |>
  arrange(meeting_date) |>
  mutate(rolling_avg = rollmean(share_climate, k = 30, fill = NA))

p <- ggplot(share, aes(x = meeting_date)) +
  geom_line(aes(y = share_climate), alpha = 0.3, color = "gray50") +
  geom_line(aes(y = rolling_avg), color = "steelblue", size = 1) +
  labs(
    title = "Share of Climate Related Debates",
    x = "Date",
    y = "Share (%)"
  ) +
  theme_minimal()
ggsave(here("Outputs/05_plot_timeseries_relative.png"))
write_rds(p, here("Outputs/05_plot_timeseries_relative.rds"))
p
```

## Climate Share Across Parties

Share of words spoken in climate-related debates per parliamentary group, measured as a percentage of each party's total word count.

```{r}
share_words <- paragraphs_classified_wide_final |>
  filter(final_climate == TRUE) 
  group_by(ParlGroupName) |>
  summarise(
    total_words = sum(paragraph_word_count),
    share = total_words / sum(total_words) * 100,
    ParlGroupName = factor(ParlGroupName, levels = rev(party_names_order))
  )

p <- share_words |>
  ggplot(aes(x = share, y = ParlGroupName, fill = ParlGroupName)) +
  geom_col() +
  scale_fill_manual(values = party_colors) +
  labs(
    title = "Climate Share of Parliamentary Debates per Party",
    x = "Share of Words in Climate Debates (%)",
    y = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none")

ggsave(here("Outputs/05_plot_climate_share_per_party.png"), width = 10, height = 6)
write_rds(p, here("Outputs/05_plot_climate_share_per_party.rds"))
p
```

## Climate Across Parties Over Time

Share of climate-related speeches per parliamentary group over time (30-session rolling average).

### Absolute terms

This means disregarding the total word count of each party. Just looking at the share of one party divided by the total word count of climate debates in that session, without accounting for the fact that some parties speak more than others.

```{r}
share <- paragraphs_classified_wide_final |>
  filter(final_climate == TRUE) |>
  group_by(meeting_date) |>
  mutate(total_climate_words = sum(paragraph_word_count)) |>
  group_by(meeting_date, ParlGroupName) |>
  summarise(
    share_party = (sum(paragraph_word_count) / first(total_climate_words)) * 100,
    .groups = "drop") |>
  mutate(
    ParlGroupName = factor(ParlGroupName, levels = party_names_order)) |>
  arrange(meeting_date, ParlGroupName) |>
  group_by(ParlGroupName) |>
  mutate(rolling_avg = rollmean(share_party, k = 30, fill = NA)) |>

p <- ggplot(share, aes(x = meeting_date, y = rolling_avg, color = ParlGroupName)) +
  geom_line(size = 1) +
  scale_color_manual(values = party_colors) +
  labs(
    title = "Absolute Share of Climate Related Debates per Party",
    x = "Date",
    y = "Share (%)",
    color = "Party Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9)
  ) +
  guides(color = guide_legend(nrow = 2))
ggsave(here("Outputs/05_plot_timeseries_per_party_absolute.png"), width = 10, height = 8)
write_rds(p, here("Outputs/05_plot_timeseries_per_party_absolute.rds"))
p
```

### relative terms

This means accounting for the total word count of each party. Looking at the share of one party in climate debates relative to that party's total word count in that session. This way we can see how much of each party's speech is dedicated to climate topics, regardless of how much they speak overall.

```{r}
share <- paragraphs_classified_wide_final |>
  group_by(meeting_date, ParlGroupName) |>
  summarise(
    total_words = sum(paragraph_word_count),
    climate_words = sum(paragraph_word_count[final_climate == TRUE]),
    .groups = "drop"
  ) |>
  mutate(
    share_party = climate_words / total_words * 100,
    ParlGroupName = factor(ParlGroupName, levels = party_names_order)
  ) |>
  arrange(meeting_date, ParlGroupName) |>
  group_by(ParlGroupName) |>
  mutate(rolling_avg = rollmean(share_party, k = 30, fill = NA))

p <- ggplot(share, aes(x = meeting_date, y = rolling_avg, color = ParlGroupName)) +
  geom_line(size = 1) +
  scale_color_manual(values = party_colors) +
  labs(
    title = "Relative Share of Climate Related Debates per Party",
    x = "Date",
    y = "Share (%)",
    color = "Party Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9)
  ) +
  guides(color = guide_legend(nrow = 2))
ggsave(here("Outputs/05_plot_timeseries_per_party_relative.png"), width = 10, height = 8)
write_rds(p, here("Outputs/05_plot_timeseries_per_party_relative.rds"))
p
```

Until around 2021, the mainstream and populist right dominated climate debates. After that point the distribution became more balanced across parties, with the Greens taking the largest share. Notably, the trajectories of the Greens and Green Liberals diverge sharply from 2020 onwards.

The graph does not account for the relative size of parliamentary groups. The mainstream left and right tend to have larger shares than the Greens and Green Liberals, with the SVP holding the largest share in parliament overall (see [parlament.ch](https://www.parlament.ch/de/organe/fraktionen/)).
