---
title: "Text Classification"
format: html
execute:
  warning: false
---

# Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
# library(openxlsx)
library(tidytext)
library(stopwords)
library(here)
library(zoo)
library(dotenv)
library(ellmer)
library(furrr)
library(future)
source(here("Scripts/00_functions.R"))
load_dot_env(here(".env"))
```


# Import Data

```{r}
#| label: data-import

paragraphs_cleaned <- read_rds(here("Data/02_paragraphs_cleaned.rds"))
# businesses_cleaned <- 
#   read_rds(here("Data/01_businesses_cleaned.rds")) |> 
#   select(BusinessShortNumber, Title)
subjects <- 
  read_rds(here("Data/02_subjects.rds")) |> 
  select(IdSubject, BusinessShortNumber, Title)
french_dictionary_curated <- 
  read_csv(here("french_dictionary_curated.csv")) |>
  rename(keyword_original = keyword_french)
german_dictionary_curated <- 
  read_csv(here("german_dictionary_curated.csv")) |>
  rename(keyword_original = keyword_german)
```

# Data Preparation

## Dictionary

Combine the curated German and French dictionaries into a single lookup table. Keywords are lowercased, stripped of non-alphanumeric characters (except hyphens and apostrophes), and whitespace-normalised.

```{r}
#| label: dictionary

dictionary_all_cleaned <- german_dictionary_curated |>
  bind_rows(french_dictionary_curated) |>
  mutate(
    keyword_original = str_to_lower(keyword_original),
    keyword_original = str_replace_all(keyword_original, "[^[:alnum:]'-]", ""),
    keyword_original = str_squish(keyword_original)
  ) |> 
  distinct(keyword_original, keyword_english)

write_rds(dictionary_all_cleaned, here("Data/05_dictionary_all_cleaned.rds"))
```

## Text Pre-Processing

Tokenise the paragraph texts into individual words, apply the same normalisation as the dictionary, and remove German and French stopwords.

```{r}
#| label: preprocessing

words <- paragraphs_cleaned |>
  select(paragraph_id, paragraph_text) |>
  unnest_tokens(
    output = "word",
    input = "paragraph_text",
    token = "words",
    drop = FALSE
  )

german_stopwords <- c(
  stopwords::stopwords("de", source = "snowball"),
  "dass"
) |>
  sort()

french_stopwords <- stopwords::stopwords("fr", source = "snowball")

stopwords_all <- tibble(stopwords = c(german_stopwords, french_stopwords))

words_cleaned <- words |>
  mutate(
    word = str_to_lower(word),
    word = str_replace_all(word, "[^[:alnum:]'-]", ""),
    word = str_squish(word)
  ) |>
  anti_join(stopwords_all, by = join_by(word == stopwords))

write_rds(words_cleaned, here("Data/05_words_cleaned.rds"))
```

# Dictionary Classification

Match each word against the dictionary and aggregate results at the paragraph level to determine which paragraphs contain at least one keyword match.

```{r}
#| label: apply-dictionary

paragraphs_classified <- words_cleaned |>
  left_join(
    dictionary_all_cleaned, 
    by = join_by(word == keyword_original), 
    relationship = "many-to-one") |>
  group_by(paragraph_id, paragraph_text) |>
  summarise(
    keywords_found = paste(na.omit(unique(keyword_english)), collapse = ", "),
    n_keywords_found = sum(!is.na(keyword_english)),
    climate_paragraph = if_else(n_keywords_found > 0, TRUE, FALSE),
    .groups = "drop"
  ) |>
  left_join(paragraphs_cleaned, by = join_by(paragraph_id))

write_rds(paragraphs_classified, here("Data/05_paragraphs_classified.rds"))
```

Do some preprocessing to prepare for final classification.

```{r}
#| label: climate-businesses

# n_distinct(subjects$IdSubject)

paragraphs_subjects_long <- left_join(
  paragraphs_classified,
  subjects,
  by = join_by(IdSubject),
  relationship = "many-to-many"
) |> 
  select(-IdSubject)

# paragraphs_subjects_long |> select(-BusinessShortNumber, -Title) |> distinct()

paragraphs_subjects_cleaned <- paragraphs_subjects_long |> 
  group_by(across(c(-BusinessShortNumber, -Title))) |> 
  summarise(
    BusinessShortNumber = paste(BusinessShortNumber, collapse = " | "),
    Title = paste(Title, collapse = " | ")
  )

# paragraphs_subjects_businesses <- left_join(
#   paragraphs_subjects,
#   businesses_cleaned,
#   by = join_by(BusinessShortNumber),
#   relationship = "many-to-one"
# )

# climate_businesses <- 
#   # tibble(
#   # climate_business = 
#     paragraphs_subjects_long |>
#     filter(climate_paragraph == TRUE) |>
#     distinct(BusinessShortNumber) |>
#     pull(BusinessShortNumber)
#   # )

paragraphs_classified_wide <- paragraphs_subjects_cleaned |>
  # mutate(
  #   climate = map_lgl(str_split(BusinessShortNumber, " \\| "), ~ any(.x %in% climate_businesses))
  # ) |>
  rename(paragraph_text = paragraph_text.x) |> 
  select(-paragraph_text.y)
```

A business is assigned a `climate_temp` label if at least 10% of its paragraphs match the dictionary or the paragraph references the climate dictionary explicitly. This threshold ensures that only businesses with substantive climate content are included.

```{r}
business_classification <- paragraphs_classified_wide |> 
  group_by(BusinessShortNumber, Title) |>
  summarise(
    n_paragraphs = n(),
    n_climate_paragraphs = sum(climate_paragraph),
    proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
    climate_temp = case_when(proportion_climate_paragraphs >= 0.1 ~ TRUE,
    .default = FALSE),
    .groups = "drop"
  )
```

In a next step, an LLM will be used to check the business Titles of climate_temp businesses and assign the final `climate` label. This is necessary to exclude false positives, i.e. businesses that contain some climate keywords but are not actually about climate topics.

```{r}
#| label: llm-classification

if (file.exists(here("Data/05_business_classification_final.rds"))) {
  
  business_classification_final <- 
    read_rds(here("Data/05_business_classification_final.rds"))
  
} else {
  
  bn_clim <- business_classification |> filter(climate_temp == TRUE)
  
  prompt <- paste(readLines(here("label_review.txt")), collapse = " ")
  
  process_text <- function(text, index) {
    chat <- chat_openai(
      model = "gpt-5-mini",
      system_prompt = prompt
    )
  
    result <- tryCatch(
      {
        chat$chat(text)
      },
      error = function(e) {
        NA_character_
      }
    )
  
    return(result)
  }
  
  plan(multisession, workers = 4)

  reviews <- future_map2_chr(
    bn_clim$Title,
    seq_len(nrow(bn_clim)),
    process_text,
    .progress = TRUE,
    .options = furrr_options(seed = TRUE)
  )

  plan(sequential)

  bn_clim_reviewed <- bn_clim |>
    mutate(llm_review = reviews)
  
  business_classification_final <- business_classification |> 
    left_join(
      bn_clim_reviewed |> select(BusinessShortNumber, llm_review),
      by = join_by(BusinessShortNumber)
    ) |> 
    mutate(
      final_climate = case_when(
        climate_temp == TRUE & llm_review == "True Positive" ~ TRUE,
        climate_temp == TRUE & llm_review == "False Positive" ~ FALSE,
        .default = FALSE
      )
    )
  
  write_rds(business_classification_final, here("Data/05_business_classification_final.rds"))
}
```

## Climate Business Analysis

Plot some summary statistics about the climate businesses
- Number of businesses in business_classification
- Number of climate_temp in business_classification
- Number of final_climate in business_classification_final

```{r}
summary_stats <- tibble(
  category = c("Total Businesses", "Climate Temp", "Final Climate"),
  count = c(
    nrow(business_classification),
    sum(business_classification$climate_temp),
    sum(business_classification_final$final_climate)
  )
)

ggplot(summary_stats, aes(x = category, y = count, fill = category)) +
  geom_col() +
  geom_text(aes(label = count), vjust = -0.5) +
  theme_minimal() +
  labs(title = "Summary of Business Classifications", x = "Category", y = "Count") +
  theme(legend.position = "none")
```

Plot the distribution of the number of pargraphs for final_climate.

```{r}
df <- business_classification_final |> 
  filter(final_climate == TRUE)

  ggplot(df, aes(x = n_paragraphs)) +
  geom_histogram() +
  theme_minimal() +
  labs(title = "Distribution of Number of Paragraphs for Final Climate Businesses", x = "Number of Paragraphs", y = "Count")
```

There are a few very large businesses that make up a large proportion of the climate discourse.

Plot the distribution of the proportion of climate paragraphs for final_climate.

```{r}
p <- plot_categorical_distribution(
  paragraphs_classified_wide_final,
  group_col = "final_climate",
  x_label = "Climate Relevant Paragraph",
  title = "Distribution of Climate Relevant Paragraphs",
  output_path = "Outputs/05_paragraphs_climate_distribution.png"
)
write_rds(p, here("Outputs/05_paragraphs_climate_distribution.rds"))
p
```

Print the largest climate relevant businesses.

```{r}
business_classification_final |> 
  filter(final_climate == TRUE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(n_paragraphs)) |> 
  slice_head(n = 10) |> 
  knitr::kable()
```

Print the largest non climate relevant businesses.

```{r}
business_classification_final |> 
  filter(final_climate == FALSE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(n_paragraphs)) |> 
  slice_head(n = 10) |> 
  knitr::kable()
```

Print the climate businesses around the mean size of a climate business in the dataset.

```{r}
business_classification_final |> 
  filter(final_climate == TRUE) |> 
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |> 
  slice_min(abs(n_paragraphs - mean(n_paragraphs)), n = 10) |> 
  knitr::kable()
```

Print the climate businesses just above the threshold.

```{r}
business_classification_final |> 
  filter(final_climate == TRUE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(proportion_climate_paragraphs) |>  
  slice_head(n = 10) |> 
  knitr::kable()
```

Few False Positives, but OK since Topic Modeling handels Outliers well. Also, very small businesses that won't have much influence on the overall topic modeling results.

Print non-climate businesses just below the threshold.

```{r}
business_classification_final |> 
  filter(final_climate == FALSE, proportion_climate_paragraphs < 0.1) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(proportion_climate_paragraphs)) |>  
  slice_head(n = 10) |> 
  knitr::kable()
```

Some medium and large businesses below the threshold with some climate relevanc, but not enough to be classified as climate relevant. This weakness is acknowlegedged, but the threshold is necessary to ensure that the final climate classification is meaningful and not diluted by businesses with only marginal climate content.

# Paragraph-Level Climate Classification

Finally, the final climate label is assigned to the paragraphs table.

```{r}
#| label: final-climate

paragraphs_classified_wide_final <- paragraphs_classified_wide |>
  left_join(
    business_classification_final |> select(BusinessShortNumber, final_climate),
    by = join_by(BusinessShortNumber)
  )

write_rds(
  paragraphs_classified_wide_final,
  here("Data/05_paragraphs_classified_wide_final.rds"))
```

# Prepare Data for Topic Modeling

Prepare climate-related paragraphs and transcripts for topic modeling. Filter out very short texts to ensure meaningful topic extraction.

```{r}
transcripts_climate_summary <- 
  paragraphs_classified_wide_final |>
  filter(final_climate == TRUE) |> 
  group_by(transcript_id) |>
  summarise(
    transcript_text = paste(paragraph_text, collapse = " "),
    transcript_word_length = str_count(transcript_text, "\\S+"),
    .groups = "drop"
  ) |>
  filter(transcript_word_length >= 50)

write_csv(
  transcripts_climate_summary,
  here("Data/05_transcripts_climate_for_topic_modeling.csv")
)
```