---
title: "Dictionary-Based Classification"
format: 
  html:
    embed-resources: true
execute:
  warning: false
---

# Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
library(tidytext)
library(stopwords)
library(here)
library(dotenv)
library(ellmer)
library(furrr)
library(future)
library(DT)
source(here("Scripts/00_functions.R"))
load_dot_env(here(".env"))
```

# Import Data

```{r}
#| label: data-import

paragraphs_cleaned <- read_csv(here("Data/02_paragraphs_cleaned.csv"))
french_dictionary_curated <- 
  read_csv(here("french_dictionary_curated.csv")) |>
  rename(keyword_original = keyword_french)
german_dictionary_curated <- 
  read_csv(here("german_dictionary_curated.csv")) |>
  rename(keyword_original = keyword_german)
```

# Data Preparation

## Dictionary

Combine the curated German and French dictionaries into a single lookup table. Keywords are lowercased, stripped of non-alphanumeric characters (except hyphens and apostrophes), and whitespace-normalised.

```{r}
#| label: dictionary

dictionary_all_cleaned <- german_dictionary_curated |>
  bind_rows(french_dictionary_curated) |>
  mutate(
    keyword_original = str_to_lower(keyword_original),
    keyword_original = str_replace_all(keyword_original, "[^[:alnum:]'-]", ""),
    keyword_original = str_squish(keyword_original)
  ) |> 
  distinct(keyword_original, keyword_english)

write_rds(dictionary_all_cleaned, here("Data/05_dictionary_all_cleaned.rds"))
```

## Text Pre-Processing

Tokenise the paragraph texts into individual words, apply the same normalisation as the dictionary, and remove German and French stopwords.

```{r}
#| label: preprocessing

words <- paragraphs_cleaned |>
  select(paragraph_id, paragraph_text) |>
  unnest_tokens(
    output = "word",
    input = "paragraph_text",
    token = "words",
    drop = FALSE
  )

german_stopwords <- c(
  stopwords::stopwords("de", source = "snowball"),
  "dass"
) |>
  sort()

french_stopwords <- stopwords::stopwords("fr", source = "snowball")

stopwords_all <- tibble(stopwords = c(german_stopwords, french_stopwords))

words_cleaned <- words |>
  mutate(
    word = str_to_lower(word),
    word = str_replace_all(word, "[^[:alnum:]'-]", ""),
    word = str_squish(word)
  ) |>
  anti_join(stopwords_all, by = join_by(word == stopwords))

write_rds(words_cleaned, here("Data/05_words_cleaned.rds"))
```

# Dictionary Classification

Match each word against the dictionary and aggregate results at the paragraph level to determine which paragraphs contain at least one keyword match.

```{r}
#| label: apply-dictionary

paragraphs_classified_dict <- words_cleaned |>
  left_join(
    dictionary_all_cleaned, 
    by = join_by(word == keyword_original), 
    relationship = "many-to-one") |>
  group_by(paragraph_id, paragraph_text) |>
  summarise(
    keywords_found = paste(na.omit(unique(keyword_english)), collapse = ", "),
    n_keywords_found = sum(!is.na(keyword_english)),
    dict_prediction = if_else(n_keywords_found > 0, TRUE, FALSE),
    .groups = "drop"
  ) |>
  left_join(paragraphs_cleaned, by = join_by(paragraph_id)) |> 
  select(-paragraph_text.y) |> 
  rename(paragraph_text = paragraph_text.x)

write_rds(paragraphs_classified_dict, here("Data/05_paragraphs_classified_dict.rds"))
```

