---
title: "Text Classification"
format: html
execute:
  warning: false
---

# Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
library(here)
source(here("Scripts/00_functions.R"))
```

# Import Data

```{r}
#| label: data-import

paragraphs_classified_wide_final <- read_rds(here("Data/05_paragraphs_classified_wide_final.rds"))
```


# Visualization Setup

Define party display order and colour palette for all subsequent plots.

```{r}
#| label: viz-setup

parlgroup_order <- c(
  "Sozialdemokratische Fraktion",
  "Grüne Fraktion",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion",
  "Grünliberale Fraktion",
  "FDP-Liberale Fraktion",
  "Fraktion der Schweizerischen Volkspartei"
)

partyabb_order <- c(
  "SP",
  "Grüne",
  "Mitte",
  "GLP",
  "FDP",
  "SVP"
)

party_colors <- c(
  "SP" = "#E4032E",
  "Grüne" = "#84B547",
  "Mitte" = "#FF8C00",
  "GLP" = "#650959ff",
  "FDP" = "#0571C1",
  "SVP" = "#00843D"
)
```

# Data Preprocessing

todo: shorten the labels (e.g. SP, Grüne, Mitte, GLP, FDP & SVP)

```{r}
visualization_data <- paragraphs_classified_wide_final |>
  mutate(
    ParlGroupName = factor(ParlGroupName, levels = rev(parlgroup_order)),
    party_abb = factor(ParlGroupName, levels = rev(parlgroup_order), labels = rev(partyabb_order))
    ) |>
  group_by(IdSession) |> 
  mutate(
    session_median_date = min(meeting_date) + as.numeric(max(meeting_date) - min(meeting_date)) / 2
  ) |> 
  ungroup()
```

# Summary Statistics

```{r}
# Summary Statistics - Overall (time-fixed)
counts <- visualization_data |>
  group_by(party_abb) |>
  summarise(
    wordcount_party = sum(paragraph_word_count),
    wordcount_party_clim = sum(paragraph_word_count[final_climate == TRUE])
  ) |>
  ungroup() |>
  mutate(
    wordcount_parl = sum(wordcount_party),
    word_count_clim = sum(wordcount_party_clim)
  )

shares <- counts |> 
  mutate(
    party_share_in_parl = wordcount_party / sum(wordcount_party) * 100,
    party_share_in_clim = wordcount_party_clim / sum(wordcount_party_clim) * 100,
    clim_share_in_party = wordcount_party_clim / wordcount_party * 100,
    avg_clim_share_in_parl = sum(wordcount_party_clim) / sum(wordcount_party) * 100,
    
    # Hypothetical climate words if each party had equal total size
    n_parties = n(),
    hypothetical_party_size = sum(wordcount_party) / n_parties,
    hypothetical_clim_words = hypothetical_party_size * (clim_share_in_party / 100),
    
    # Size-adjusted climate share (this will sum to 100%)
    size_adjusted_clim_share = hypothetical_clim_words / sum(hypothetical_clim_words) * 100,
    
    # Expected climate words if party spoke at parliament average rate
    expected_clim_words = wordcount_party * (avg_clim_share_in_parl / 100),
    
    # Ratio of actual to expected (>1 means above average, <1 below)
    clim_intensity_ratio = wordcount_party_clim / expected_clim_words
  )

# Summary Statistics - By Session (time-variable)
session_counts <- visualization_data |> 
  group_by(IdSession, session_median_date, party_abb) |>
  summarise(
    wordcount_party = sum(paragraph_word_count),
    wordcount_party_clim = sum(paragraph_word_count[final_climate == TRUE]),
    .groups = "drop"
  ) |>
  group_by(IdSession, session_median_date) |>
  mutate(
    session_wordcount = sum(wordcount_party),
    session_wordcount_clim = sum(wordcount_party_clim)
  ) |>
  ungroup()

session_shares <- session_counts |> 
  mutate(
    party_share_in_parl = wordcount_party / session_wordcount * 100,
    party_share_in_clim = wordcount_party_clim / session_wordcount_clim * 100,
    clim_share_in_party = wordcount_party_clim / wordcount_party * 100,
    avg_clim_share_in_session = session_wordcount_clim / session_wordcount * 100,
    
    # Hypothetical climate words if each party had equal total size
    n_parties = n_distinct(party_abb),
    hypothetical_party_size = session_wordcount / n_parties,
    hypothetical_clim_words = hypothetical_party_size * (clim_share_in_party / 100),
    
    # Size-adjusted climate share
    size_adjusted_clim_share = hypothetical_clim_words / sum(hypothetical_clim_words) * 100,
    
    # Expected climate words if party spoke at parliament average rate
    expected_clim_words = wordcount_party * (avg_clim_share_in_session / 100),
    
    # Ratio of actual to expected
    clim_intensity_ratio = wordcount_party_clim / expected_clim_words
  )
```

# Climate Discourse Over Time

todo: Sondersessionen rausnehmen? Oder zumindest markieren? Weil die nicht mit regulären Sessions vergleichbar sind. Oder wir nehmen den Teil ganz raus weil nicht relevant.

```{r}
session_climate_share <- visualization_data |>
  group_by(IdSession, session_median_date) |>
  summarise(
    share_climate = sum(paragraph_word_count[final_climate == TRUE]) / sum(paragraph_word_count) * 100,
    .groups = "drop"
  )

p <- ggplot(session_climate_share, aes(x = session_median_date, y = share_climate)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "steelblue", size = 2) +
  labs(
    title = "Share of Climate Related Debates",
    x = "Session Date",
    y = "Share (%)"
  ) +
  theme_minimal()
ggsave(here("Outputs/05_plot_timeseries.png"))
write_rds(p, here("Outputs/05_plot_timeseries.rds"))
p
```

```{r}
# # Insepect businesses of a single session
# paragraphs_classified_wide_final |> 
#   filter(IdSession == 5116) |> 
#   distinct(Title, final_climate) |>
#   arrange(final_climate) |> 
#   print(n = "Inf")
```

## Climate Discourse Across Parties

### Time-Fixed

First Question: How much of each party's speech time do the parties devote to climate change? The grey line shows the parliament average at `r unique(shares$avg_clim_share_in_parl)`. In white are the percentages of climate speech in comparison to their total speech time.

```{r}
p <- shares |>
  ggplot(aes(x = clim_intensity_ratio, y = party_abb, fill = party_abb)) +
  geom_col() +
  scale_fill_manual(values = party_colors) +
  geom_text(aes(label = ifelse(clim_share_in_party > 3, 
                                sprintf("%.1f%%", clim_share_in_party), 
                                "")),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold",
            size = 3.5) +
  labs(
    title = "Contribution to Climate Debate per Party",
    x = "Actual Climate Contribution Compared to Parliament Average",
    y = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50")

ggsave(here("Outputs/05_plot_climate_factor_per_party.png"))
write_rds(p, here("Outputs/05_plot_climate_factor_per_party.rds"))
p
```

Second Question: Which parties dominate the climate discourse? I answer this question by showing each parties contribution to the total climate discourse. The second bar shows how the climate discourse would like if each party were of equal size.

```{r}

# Reshape data for two bars
shares_long <- shares |>
  select(party_abb, size_adjusted_clim_share, party_share_in_clim) |>
  pivot_longer(
    cols = c(size_adjusted_clim_share, party_share_in_clim),
    names_to = "measure",
    values_to = "percentage"
  ) |>
  mutate(
    measure = factor(measure, 
                    levels = c("size_adjusted_clim_share", "party_share_in_clim"),
                    labels = c("Size-Adjusted", "Actual"))
  )

p <- shares_long |>
  ggplot(aes(x = percentage, y = measure, fill = party_abb)) +
  geom_col(position = "stack", width = 0.6) +
  geom_text(aes(label = ifelse(percentage > 3, 
                                sprintf("%.1f%%", percentage), 
                                "")),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold",
            size = 3.5) +
  scale_fill_manual(values = party_colors) +
  labs(
    title = "Climate Share of Parliamentary Debates per Party",
    subtitle = "Size-Adjusted: If all parties had equal speaking time | Actual: Real distribution",
    x = "Share of Words in Climate Debates (%)",
    y = NULL,
    fill = "Party"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )

ggsave(here("Outputs/05_plot_climate_share_per_party.png"))
write_rds(p, here("Outputs/05_plot_climate_share_per_party.rds"))
p
```

todo: add a third bar that shows share of words in all debates!

### Time Variable

Share of climate-related speeches per parliamentary group over time (30-days rolling average).

First plot:

This means disregarding the total word count of each party. Just looking at the share of one party divided by the total word count of the climate debate in that session, without accounting for the fact that some parties speak more than others.

```{r}
p <- session_shares |> 
  ggplot(aes(x = session_median_date, y = party_share_in_clim, color = party_abb)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  scale_color_manual(values = party_colors) +
  labs(
    title = "Party Share in Climate Discourse",
    x = "Session Date",
    y = "Share (%)",
    color = "Party Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
ggsave(here("Outputs/05_plot_climshare_timeseries.png"))
write_rds(p, here("Outputs/05_plot_climshare_timeseries.rds"))
p
```

Until around 2021, the mainstream and populist right dominated climate debates. After that point the distribution became more balanced across parties, with the Greens taking the largest share. Notably, the trajectories of the Greens and Green Liberals diverge sharply from 2020 onwards.

The graph does not account for the relative size of parliamentary groups. The mainstream left and right tend to have larger shares than the Greens and Green Liberals, with the SVP holding the largest share in parliament overall (see [parlament.ch](https://www.parlament.ch/de/organe/fraktionen/)).

Second Plot:

```{r}
p <- session_shares |>
  ggplot(aes(x = session_median_date, y = clim_share_in_party, color = party_abb)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  scale_color_manual(values = party_colors) +
  labs(
    title = "Climate Share in Party's Speeches",
    x = "Session Date",
    y = "Share (%)",
    color = "Party Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
ggsave(here("Outputs/05_plot_partyshare_timeseries.png"))
write_rds(p, here("Outputs/05_plot_partyshare_timeseries.rds"))
p
```

todo: überlegen ob nur reguläre Sessionen genommen werden sollen für die Vergleichbarkeit? Oder wir nehmen den Teil ganz raus weil nicht relevant.

When we look at the contribution of a single party to the climate discourse, we see that the dominating factor is the presence of the climate in the parliament. When climate is the topic, every party contributes about the same with more variation happening towards the end of the 10-year period. This suggests that the differences in absolute share are largely driven by the overall presence of climate topics in the parliament, rather than by specific parties dominating the climate debate when it occurs.

