---
title: "Text Classification"
format: 
  html:
    embed-resources: true
execute:
  warning: false
---

# Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
library(tidytext)
library(stopwords)
library(here)
library(dotenv)
library(ellmer)
library(furrr)
library(future)
library(DT)
source(here("Scripts/00_functions.R"))
load_dot_env(here(".env"))
```

# Import Data

```{r}
subjects <- 
  read_rds(here("Data/02_subjects.rds")) |> 
  select(IdSubject, BusinessShortNumber, Title)
```

# Dictionary-Based Classification

Here the dictionary-based classification results will be imported. The classification is based on the presence of keywords from the curated German and French dictionaries in the paragraphs. A paragraph is classified as climate-relevant if it contains at least one keyword from the dictionary.

```{r}
#| label: dict-classification

paragraphs_classified_dict <- read_rds(here("Data/05_paragraphs_classified_dict.rds")) |> 
  mutate(IdSubject = as.character(IdSubject))
```

# BERT Classification

Here the BERT-based classification results will be imported. 

```{r}
#| label: bert-classification

classified_paragraphs_bert <- read_csv(here("Model-Outputs/classified_paragraphs.csv"))
classified_paragraphs_bert_cleaned <- classified_paragraphs_bert |>
  mutate(predicted_label = ifelse(predicted_label == 1, TRUE, FALSE)) |> 
  select(-token_length) |> 
  rename(paragraph_id = id)
```

# Compare bert- and dict-based classification

Compare the dictionary-based classification with the BERT-based classification. 

```{r}
paragraphs_classified_joint <- paragraphs_classified_dict |> 
  left_join(
    classified_paragraphs_bert_cleaned |> select(paragraph_id, predicted_label),
    by = join_by(paragraph_id)
  ) |>
  drop_na(predicted_label)   
  # mutate(
  #   either_prediction = dict_prediction | predicted_label,
  #   both_predictions = dict_prediction & predicted_label)
```

Visualize the comparison results.

Show the dict prediction, the bert prediction and the overlap between the two predictions.

```{r}
df_plot <- paragraphs_classified_joint |> 
  summarise(
    Dictionary = sum(dict_prediction),
    Transformer = sum(predicted_label),
    Overlap = sum(dict_prediction & predicted_label)
  ) |> 
  pivot_longer(cols = everything(), names_to = "Category", values_to = "Count")

p <- ggplot(df_plot, aes(x = Category, y = Count, fill = Category)) +
  geom_col() +
  geom_text(aes(label = Count), vjust = 2.0) +
  theme_minimal() +
  labs(x = "Prediction Category", y = "Count") +
  theme(legend.position = "none")
ggsave(here("Outputs/05_prediction_comparison.png"), p)
write_rds(p, here("Outputs/05_prediction_comparison.rds"))
p
```

# Transcript Classification

```{r}
# transcript_classification <- paragraphs_classified_joint |> 
#   group_by(transcript_id) |>
#   summarise(
#     n_paragraphs = n(),
#     n_climate_paragraphs = sum(predicted_label),
#     proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
#     climate_transcript = case_when(n_climate_paragraphs > 0 ~ TRUE,
#     .default = FALSE),
#     transcript_text = paste(paragraph_text, collapse = " "),
#     transcript_word_length = str_count(transcript_text, "\\S+"),
#     .groups = "drop"
#   ) |> 
#   filter(climate_transcript == TRUE, transcript_word_length >= 50)

# write_csv(
#   transcript_classification,
#   here("Data/05_transcripts_climate.csv")
# )
```

# Pre-Processing for Business Classification

Do some preprocessing to prepare for final classification.

```{r}
# n_distinct(subjects$IdSubject)

# Create long table: paragraphs can have more than one bsn assigned

paragraphs_subjects_long <- left_join(
  paragraphs_classified_dict,
  subjects,
  by = join_by(IdSubject),
  relationship = "many-to-many"
) |> 
  select(-IdSubject)

paragraphs_classified_wide <- paragraphs_subjects_long |> 
  group_by(across(c(-BusinessShortNumber, -Title))) |> 
  summarise(
    BusinessShortNumber = paste(BusinessShortNumber, collapse = " | "),
    Title = paste(Title, collapse = " | "),
    .groups = "drop"
  )
```

# Sensitivity Checks for Threshold Selection

```{r}
# create variables with different thresholds for the proportion of climate paragraphs to be classified as climate relevant. This will be used for sensitivity checks later on.

# thresholds <- c(0.03, 0.05, 0.08, 0.1, 0.12, 0.15)

business_classification <- paragraphs_classified_wide |>
  group_by(BusinessShortNumber, Title) |>
  summarise(
    n_paragraphs = n(),
    n_climate_paragraphs = sum(dict_prediction),
    proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
    threshold_03 = ifelse(proportion_climate_paragraphs >= 0.03, TRUE, FALSE),
    threshold_05 = ifelse(proportion_climate_paragraphs >= 0.05, TRUE, FALSE),
    threshold_08 = ifelse(proportion_climate_paragraphs >= 0.08, TRUE, FALSE),
    threshold_10 = ifelse(proportion_climate_paragraphs >= 0.1, TRUE, FALSE),
    threshold_12 = ifelse(proportion_climate_paragraphs >= 0.12, TRUE, FALSE),
    threshold_15 = ifelse(proportion_climate_paragraphs >= 0.15, TRUE, FALSE),
    .groups = "drop"
  )

# label the paragraphs table with each threshold
# pseudo code:
# when BusinesShortNumber in threshold_03 == TRUE then threshold_03_paragraph = TRUE else FALSE

paragraphs_classified_sensitivity <- paragraphs_classified_wide |>
  left_join(
    business_classification |> select(BusinessShortNumber, starts_with("threshold")),
    by = join_by(BusinessShortNumber)
  ) |>
  group_by(paragraph_id) |>
  summarise(
    across(starts_with("threshold"), ~ as.integer(any(. == TRUE))),
    .groups = "drop"
  ) |>
  summarise(across(starts_with("threshold"), sum)) |>
  pivot_longer(cols = everything(), names_to = "threshold", values_to = "n_paragraphs_classified")


# plot the number of paragraphs for each threshold
p <- ggplot(paragraphs_classified_sensitivity, aes(x = threshold, y = n_paragraphs_classified, fill = threshold)) +
  geom_col() +
  geom_text(aes(label = n_paragraphs_classified), vjust = 2.0) +
  theme_minimal() +
  labs(x = "Proportion of Climate Paragraphs Threshold", y = "Count of Climate Paragraphs") +
  theme(legend.position = "none")
ggsave(here("Outputs/05_threshold_counts.png"), p)
write_rds(p, here("Outputs/05_threshold_counts.rds"))
p
```

Plot a histogram of the distribution of the proportion of climate paragraphs for all businesses.

```{r}
#| label: histogram

df_plot <- business_classification |> 
  filter(proportion_climate_paragraphs >= 0.01 & proportion_climate_paragraphs <= 0.2)

p <- ggplot(df_plot, aes(x = proportion_climate_paragraphs)) +
  geom_histogram() +
  theme_minimal() +
  labs(
    x = "Proportion of Climate Paragraphs", 
    y = "Count"
  )
ggsave(here("Outputs/05_proportion_climate_paragraphs_histogram.png"), p)
write_rds(p, here("Outputs/05_proportion_climate_paragraphs_histogram.rds"))
p
```

# Temporary business classification

A business is assigned a `climate_temp` label if at least 10% of its paragraphs match the dictionary or the paragraph references the climate dictionary explicitly. This threshold ensures that only businesses with substantive climate content are included. The threshold was determined by manually inspecting the distribution of the proportion of climate paragraphs across businesses and identifying a natural cutoff point that balances inclusivity with relevance.

```{r}
#| label: business-classification

final_threshold <- 0.1

business_classification <- paragraphs_classified_wide |> 
  group_by(BusinessShortNumber, Title) |>
  summarise(
    n_paragraphs = n(),
    n_climate_paragraphs = sum(dict_prediction),
    proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
    climate_temp = case_when(proportion_climate_paragraphs >= final_threshold ~ TRUE,
    .default = FALSE),
    large_bsn_check = case_when(n_climate_paragraphs > 1 & n_paragraphs >= 30 ~ TRUE,
    .default = FALSE),
    .groups = "drop"
  )
```


# LLM Review

In a next step, an LLM will be used to check the business Titles of climate_temp businesses and assign the final `climate` label. This is necessary to exclude false positives, i.e. businesses that contain some climate keywords but are not actually about climate topics.

```{r}
#| label: llm-classification

if (file.exists(here("Data/05_business_classification_final.rds"))) {
  
  business_classification_final <- 
    read_rds(here("Data/05_business_classification_final.rds"))
  
} else {
  
  bn_clim <- business_classification |> filter(climate_temp == TRUE)
  
  prompt <- paste(readLines(here("label_review.txt")), collapse = " ")
  
  process_text <- function(text, index) {
    chat <- chat_openai(
      model = "gpt-5-mini",
      system_prompt = prompt
    )
  
    result <- tryCatch(
      {
        chat$chat(text)
      },
      error = function(e) {
        NA_character_
      }
    )
  
    return(result)
  }
  
  plan(multisession, workers = 4)

  reviews <- future_map2_chr(
    bn_clim$Title,
    seq_len(nrow(bn_clim)),
    process_text,
    .progress = TRUE,
    .options = furrr_options(seed = TRUE)
  )

  plan(sequential)

  bn_clim_reviewed <- bn_clim |>
    mutate(llm_review = reviews)
  
  business_classification_final <- business_classification |> 
    left_join(
      bn_clim_reviewed |> select(BusinessShortNumber, llm_review),
      by = join_by(BusinessShortNumber)
    ) |> 
    mutate(
      final_climate = case_when(
        climate_temp == TRUE & llm_review == "True Positive" ~ TRUE,
        climate_temp == TRUE & llm_review == "False Positive" ~ FALSE,
        .default = FALSE
      )
    )
  
  write_rds(business_classification_final, here("Data/05_business_classification_final.rds"))
}
```

# Climate Business Analysis

Plot some summary statistics about the climate businesses
- Number of businesses with proportion_climate_paragraphs > 0
- Number of businesses in climate_temp
- Number of businesses in final_climate

```{r}
categories = c("Dictionary Match", ">10% Threshold", "After LLM Review")

summary_stats <- tibble(
  category = factor(categories, levels = categories),
  count = c(
    nrow(business_classification_final |> filter(proportion_climate_paragraphs > 0)),
    sum(business_classification_final$climate_temp),
    sum(business_classification_final$final_climate)
  )
)

p <- ggplot(summary_stats, aes(x = category, y = count, fill = category)) +
  geom_col() +
  geom_text(aes(label = count), vjust = 1.5) +
  theme_minimal() +
  labs(title = "Summary of Business Classifications", x = "Category", y = "Count") +
  theme(legend.position = "none")
ggsave(here("Outputs/05_business_classification_summary.png"), p)
write_rds(p, here("Outputs/05_business_classification_summary.rds"))
p
```

Plot the distribution of the number of pargraphs for final_climate.

```{r}
df <- business_classification_final |> 
  filter(final_climate == TRUE)

ggplot(df, aes(x = n_paragraphs)) +
geom_histogram() +
theme_minimal() +
labs(
  title = "Distribution of Number of Paragraphs for Final Climate Businesses", 
  x = "Number of Paragraphs", 
  y = "Count"
)
```

There are a few very large businesses that make up a large proportion of the climate discourse.

Print the largest climate relevant businesses.

```{r}
business_classification_final |> 
  filter(final_climate == TRUE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(n_paragraphs)) |> 
  slice_head(n = 10) |> 
  datatable()
```

Print the largest non climate relevant businesses.

```{r}
business_classification_final |> 
  filter(final_climate == FALSE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(n_paragraphs)) |> 
  slice_head(n = 10) |> 
  datatable()
```

Print the climate businesses around the mean size of a climate business in the dataset.

```{r}
business_classification_final |> 
  filter(final_climate == TRUE) |> 
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |> 
  slice_min(abs(n_paragraphs - mean(n_paragraphs)), n = 10) |> 
  datatable()
```

Print the climate businesses just above the threshold.

```{r}
business_classification_final |> 
  filter(final_climate == TRUE) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(proportion_climate_paragraphs) |>  
  slice_head(n = 10) |> 
  datatable()
```

Still a few False Positives, but OK since Topic Modeling handels Outliers well. Also, very small businesses that won't have much influence on the overall topic modeling results.

Print non-climate businesses just below the threshold.

```{r}
business_classification_final |> 
  filter(final_climate == FALSE, proportion_climate_paragraphs < final_threshold) |>
  distinct(Title, n_paragraphs, proportion_climate_paragraphs) |>
  arrange(desc(proportion_climate_paragraphs)) |>  
  slice_head(n = 10) |> 
  datatable()
```

Some medium-size businesses below the threshold with some climate relevance, but not enough to be classified as climate relevant. This weakness is acknowleged, but the threshold is necessary to ensure that the final climate classification is meaningful and not diluted by businesses with only marginal climate content.


# Paragraph-Level Climate Classification

Finally, the final climate label is assigned to the paragraphs table.

```{r}
#| label: final-climate

paragraphs_classified_wide_final <- paragraphs_classified_wide |>
  left_join(
    business_classification_final |> select(BusinessShortNumber, final_climate),
    by = join_by(BusinessShortNumber)
  )

write_rds(
  paragraphs_classified_wide_final,
  here("Data/05_paragraphs_classified_wide_final.rds"))
```

Plot the distribution of the proportion of climate paragraphs for final_climate.

```{r}
p <- plot_categorical_distribution(
  paragraphs_classified_wide_final,
  group_col = "final_climate",
  x_label = "Climate Relevant Paragraph",
  title = "Distribution of Climate Relevant Paragraphs",
  output_path = "Outputs/05_paragraphs_climate_distribution.png"
)
write_rds(p, here("Outputs/05_paragraphs_climate_distribution.rds"))
p
```



# Prepare Data for Topic Modeling

Prepare climate-related paragraphs and transcripts for topic modeling. Filter out very short texts to ensure meaningful topic extraction.

```{r}
transcripts_climate_summary <- 
  paragraphs_classified_wide_final |>
  filter(final_climate == TRUE) |> 
  group_by(transcript_id) |>
  summarise(
    transcript_text = paste(paragraph_text, collapse = " "),
    transcript_word_length = str_count(transcript_text, "\\S+"),
    .groups = "drop"
  ) |>
  filter(transcript_word_length >= 50)
  
write_csv(
  transcripts_climate_summary,
  here("Data/05_transcripts_climate_for_topic_modeling.csv")
)
```