---
title: "Topic Analysis"
format: html
execute:
  warning: false
---

# Setup

Import packages and helper functions.

```{r}
#| label: setup

library(tidyverse)
library(here)
library(knitr)
source(here("Scripts/00_functions.R"))
```

# Import Data

Read topic modeling results and join with paragraph-level data.

```{r}
#| label: import-data

documents_with_topics <- read_csv(here("Data/06_documents_with_topics.csv")) |>
  mutate(
    transcript_id = as.character(transcript_id),
    topic = sprintf("%02d", as.numeric(topic))
  ) |>
  filter(topic != "-1")

topic_info <- read_csv(here("Data/06_topic_info.csv")) |>
  rename(topic = Topic, topic_name = Name) |>
  mutate(topic = sprintf("%02d", as.numeric(topic))) |>
  filter(topic != "-1")

paragraphs_climate <- read_rds(here("Data/05_paragraphs_classified_wide_final.rds")) |>
  filter(final_climate == TRUE) |>
  mutate(
    ParlGroupName = case_when(
      ParlGroupAbbreviation %in% c("M-E", "BD") ~ "Die Mitte-Fraktion / CVP-& BD-Fraktion",
      .default = ParlGroupName
    )
  )
```

# Visualization Setup

Define party display order and colour palette for all subsequent plots.

```{r}
#| label: viz-setup

party_names_order <- c(
  "Sozialdemokratische Fraktion",
  "Grüne Fraktion",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion",
  "Grünliberale Fraktion",
  "FDP-Liberale Fraktion",
  "Fraktion der Schweizerischen Volkspartei"
)

party_abb_order <- c("SP", "Grüne", "Mitte", "GLP", "FDP", "SVP")

mapping_tbl <- tibble(
  ParlGroupName = party_names_order,
  party_abb = party_abb_order
)

party_colors <- c(
  "Sozialdemokratische Fraktion" = "#E4032E",
  "Grüne Fraktion" = "#84B547",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion" = "#FF8C00",
  "Grünliberale Fraktion" = "#650959ff",
  "FDP-Liberale Fraktion" = "#0571C1",
  "Fraktion der Schweizerischen Volkspartei" = "#00843D"
)
```

# Data Preparation

Join topic assignments with paragraph-level metadata and create visualization data.

```{r}
#| label: data-prep

documents_with_topics_wide <- documents_with_topics |>
  inner_join(topic_info, by = join_by(topic))

paragraphs_topics_wide <- documents_with_topics_wide |>
  inner_join(
    paragraphs_climate,
    by = join_by(transcript_id),
    relationship = "one-to-many"
  ) |>
  drop_na(ParlGroupName) |>
  mutate(
    ParlGroupName = factor(ParlGroupName, levels = rev(party_names_order)),
    topic = factor(topic, levels = rev(sort(unique(topic))))
  ) |>
  # left_join(mapping_tbl) |>
  # mutate(party_abb = factor(party_abb, levels = rev(party_abb_order))) |>
  group_by(IdSession) |> 
  mutate(
    session_median_date = min(meeting_date) + as.numeric(max(meeting_date) - min(meeting_date)) / 2
  ) |> 
  ungroup()
```

# Topic Representation

Display topic labels generated by the representation models.

```{r}
#| label: topic-representation

topic_representation <- topic_info |>
  select(topic, OpenAI)

write_rds(topic_representation, here("Outputs/07_topic_representation.rds"))

topic_representation |>
  kable(
    caption = "Topic Representation based on OpenAI summaries",
    col.names = c("Topic", "Representation")
  )
```

# Topic Distribution

## Overall Distribution

Overall distribution of word counts across topics.

```{r}
#| label: topic-distribution

df_topic_distribution <- paragraphs_topics_wide |>
  group_by(topic) |>
  summarise(total_words = sum(paragraph_word_count)) |>
  mutate(
    pct_total = total_words / sum(total_words) * 100
  )

p <- ggplot(df_topic_distribution, aes(x = topic, y = pct_total)) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = paste0(round(pct_total, 1), "%")),
    hjust = -0.2, 
    size = 3
  ) +
  coord_flip() +
  labs(
    x = "Topic",
    y = "Percentage of Total Words",
    title = "Distribution of Words across Topics"
  ) +
  theme_minimal()

ggsave(here("Outputs/07_plot_topic_distribution.png"))
write_rds(p, here("Outputs/07_plot_topic_distribution.rds"))
p
```

## Party Distribution per Topic

Party composition across all topics (stacked bar charts) based on word counts.

```{r}
#| label: party-distribution-per-topic

df_plot <- paragraphs_topics_wide |>
  group_by(topic, ParlGroupName) |>
  summarise(
    group_words = sum(paragraph_word_count),
    .groups = 'drop_last'
  ) |>
  mutate(
    pct_of_topic = group_words / sum(group_words) * 100
  ) |>
  drop_na()

p <- ggplot(
  df_plot,
  aes(x = pct_of_topic, y = topic, fill = ParlGroupName)
) +
  geom_col(position = "stack") +
  geom_text(aes(label = ifelse(pct_of_topic > 15,
                                sprintf("%.1f%%", pct_of_topic),
                                "")),
            position = position_stack(vjust = 0.5),
            color = "white",
            size = 3.5) +
  scale_fill_manual(values = party_colors) +
  labs(
    title = "Party Share per Topic",
    x = "Share (%)",
    y = "Topic",
    fill = "Party"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5)
  )

ggsave(here("Outputs/07_plot_topic_party_distribution.png"), width = 10, height = 6)
write_rds(p, here("Outputs/07_plot_topic_party_distribution.rds"))
p
```

# Topic Preferences by Party

Calculate which parties over- or under-emphasize specific topics relative to their overall speaking time.

```{r}
#| label: topic-preferences

# Calculate overall speech proportion for each party based on word counts
party_weights <- paragraphs_topics_wide |>
  group_by(ParlGroupName) |>
  summarise(
    total_party_words = sum(paragraph_word_count),
    .groups = 'drop'
  ) |>
  mutate(
    party_proportion = total_party_words / sum(total_party_words)
  )

# Calculate topic preferences
topic_preferences <- paragraphs_topics_wide |>
  group_by(topic, ParlGroupName) |>
  summarise(
    party_topic_words = sum(paragraph_word_count),
    .groups = 'drop'
  ) |>
  # Add party weights
  left_join(party_weights, by = "ParlGroupName") |>
  # Calculate weighted mean for each topic
  group_by(topic) |>
  mutate(
    mean_topic_words = sum(party_topic_words) / n_distinct(ParlGroupName),
    weighted_mean_topic = mean_topic_words * party_proportion,
    topic_factor = party_topic_words / weighted_mean_topic
  ) |>
  ungroup() |>
  select(topic, ParlGroupName, party_topic_words, weighted_mean_topic, topic_factor, party_proportion) |>
  arrange(topic, desc(topic_factor))

# Create a wide table for better visualization
topic_preferences_wide <- topic_preferences |>
  select(topic, ParlGroupName, topic_factor) |>
  pivot_wider(
    names_from = ParlGroupName,
    values_from = topic_factor
  )

write_rds(topic_preferences, here("Outputs/07_topic_preferences.rds"))
write_rds(topic_preferences_wide, here("Outputs/07_topic_preferences_wide.rds"))

topic_preferences_wide |>
  kable(
    caption = "Topic Preference Factors by Party (>1 = over-representation, <1 = under-representation)",
    digits = 2
  )
```

## Visualize Topic Preferences

Heatmap showing which parties emphasize which topics.

```{r}
#| label: topic-preferences-heatmap

p <- ggplot(topic_preferences, aes(x = ParlGroupName, y = topic, fill = topic_factor)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(topic_factor, 2)), size = 3, color = "white") +
  scale_fill_gradient2(
    low = "blue", 
    mid = "white", 
    high = "red", 
    midpoint = 1,
    name = "Preference\nFactor"
  ) +
  labs(
    title = "Topic Preferences by Party",
    subtitle = "Values >1 indicate over-representation, <1 under-representation",
    x = NULL,
    y = "Topic"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

ggsave(here("Outputs/07_plot_topic_preferences_heatmap.png"), width = 12, height = 8)
write_rds(p, here("Outputs/07_plot_topic_preferences_heatmap.rds"))
p
```

# Topic Trends Over Time

Analyze how topic prevalence changes over parliamentary sessions based on word counts.

```{r}
#| label: topic-trends

# Calculate topic share by session using word counts
session_topic_shares <- paragraphs_topics_wide |>
  group_by(IdSession, session_median_date, topic) |>
  summarise(
    topic_words = sum(paragraph_word_count),
    .groups = 'drop'
  ) |>
  group_by(IdSession, session_median_date) |>
  mutate(
    total_words = sum(topic_words),
    topic_share = topic_words / total_words * 100
  ) |>
  ungroup()

# Plot top 5 topics over time
top_topics <- df_topic_distribution |>
  arrange(desc(total_words)) |>
  head(5) |>
  pull(topic)

p <- session_topic_shares |>
  filter(topic %in% top_topics) |>
  ggplot(aes(x = session_median_date, y = topic_share, color = topic)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  labs(
    title = "Top 5 Topics Over Time",
    x = "Session Date",
    y = "Share of Climate Discourse (%)",
    color = "Topic"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"
  )

ggsave(here("Outputs/07_plot_topic_trends.png"), width = 10, height = 6)
write_rds(p, here("Outputs/07_plot_topic_trends.rds"))
p
```

# Party Topic Focus Over Time

Track how each party's topic focus evolves across sessions based on word counts.

```{r}
#| label: party-topic-time

# Calculate party-specific topic shares by session using word counts
party_session_topics <- paragraphs_topics_wide |>
  group_by(IdSession, session_median_date, ParlGroupName, topic) |>
  summarise(
    topic_words = sum(paragraph_word_count),
    .groups = 'drop'
  ) |>
  group_by(IdSession, session_median_date, ParlGroupName) |>
  mutate(
    party_total = sum(topic_words),
    topic_share_in_party = topic_words / party_total * 100
  ) |>
  ungroup()

# Select top topic for visualization
selected_topic <- top_topics[1]

p <- party_session_topics |>
  filter(topic == selected_topic) |>
  ggplot(aes(x = session_median_date, y = topic_share_in_party, color = ParlGroupName)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  scale_color_manual(values = party_colors) +
  labs(
    title = paste("Party Focus on Topic", selected_topic, "Over Time"),
    x = "Session Date",
    y = "Share of Party's Climate Discourse (%)",
    color = "Party"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"
  )

ggsave(here("Outputs/07_plot_party_topic_trends.png"), width = 10, height = 6)
write_rds(p, here("Outputs/07_plot_party_topic_trends.rds"))
p
```
