---
title: "Topic Analysis"
format: html
execute:
  warning: false
---

# Setup

Import packages and helper functions.

```{r}
#| label: setup

library(tidyverse)
library(tidytext)
library(here)
library(knitr)
library(DT)
library(purrr)
source(here("Scripts/00_functions.R"))
```

# Import Data

Read topic modeling results and join with paragraph-level data.

```{r}
#| label: import-data

documents_with_topics <- read_csv(here("Data/06_documents_with_topics.csv")) |>
  # filter(topic != "-1") |> 
  mutate(
    transcript_id = as.character(transcript_id),
    topic_id = topic + 1
  ) |> 
  select(-topic)

topic_info <- read_csv(here("Data/06_topic_info.csv")) |>
  # filter(topic != "-1") |> 
  # rename(topic = Topic, topic_name = Name) |>
  mutate(
    # topic = sprintf("%02d", as.numeric(topic)),
    topic_id = row_number() - 1, 
    OpenAI = str_remove_all(OpenAI, "\\[|\\]|'"),
    OpenAI = ifelse(topic_id == 0, "Outliers", OpenAI),
    KeyBERT = str_remove_all(KeyBERT, "\\[|\\]|'"),
    MMR = str_remove_all(MMR, "\\[|\\]|'")
  ) |> 
  select(topic_id, everything(), -c(Topic, Name, Representation))

write_rds(topic_info, here("Data/06_r_topic_info.rds"))  

paragraphs_climate <- read_rds(here("Data/05_paragraphs_classified_wide_final.rds")) |>
  filter(final_climate == TRUE)
```

# Visualization Setup

Define party display order and colour palette for all subsequent plots.

```{r}
#| label: viz-setup

parlgroup_order <- c(
  "Sozialdemokratische Fraktion",
  "Grüne Fraktion",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion",
  "Grünliberale Fraktion",
  "FDP-Liberale Fraktion",
  "Fraktion der Schweizerischen Volkspartei"
)

partyabb_order <- c(
  "SP",
  "Grüne",
  "Mitte",
  "GLP",
  "FDP",
  "SVP"
)

party_colors <- c(
  "SP" = "#E4032E",
  "Grüne" = "#84B547",
  "Mitte" = "#FF8C00",
  "GLP" = "#650959ff",
  "FDP" = "#0571C1",
  "SVP" = "#00843D"
)
```

# Data Preparation

Join topic assignments with paragraph-level metadata and create visualization data.

```{r}
#| label: data-prep

documents_with_topics_wide <- documents_with_topics |>
  inner_join(topic_info, by = join_by(topic_id))

paragraphs_topics_wide <- documents_with_topics_wide |>
  inner_join(
    paragraphs_climate,
    by = join_by(transcript_id),
    relationship = "one-to-many"
  ) |>
  drop_na(ParlGroupName) |>
  mutate(
    ParlGroupName = factor(ParlGroupName, levels = rev(parlgroup_order)),
    party_abb = factor(ParlGroupName, levels = rev(parlgroup_order), labels = rev(partyabb_order))#,
    # topic = factor(topic, levels = rev(sort(unique(topic))))
  ) |>
  # left_join(mapping_tbl) |>
  # mutate(party_abb = factor(party_abb, levels = rev(party_abb_order))) |>
  group_by(IdSession) |> 
  mutate(
    session_median_date = min(meeting_date) + as.numeric(max(meeting_date) - min(meeting_date)) / 2
  ) |> 
  ungroup()
```

# Topic Representation

Display topic labels generated by the representation models.

```{r}
#| label: topic-representation

topic_overview <- 
  paragraphs_topics_wide |> 
  # filter(topic != "-1") |>
  group_by(topic_id, OpenAI) |> 
  summarise(
    count = n_distinct(transcript_id),
    .groups = "drop"
  ) |> 
  mutate(pct_total = round(count / sum(count) * 100, 1)) |> 
  arrange(topic_id)

write_rds(topic_overview, here("Outputs/06_tbl_topic_overview.rds"))
topic_overview |> 
  datatable(colnames = c("ID", "LLM Name", "Count", "%"))
```

Display KeyBERT and MMR representations for each topic.

```{r}
representation_overview <- topic_info |> 
  select(topic_id, OpenAI, KeyBERT, MMR)

representation_overview |> 
  datatable()
```

```{r}
#| label: topic-distribution

df_topic_distribution <- paragraphs_topics_wide |>
  group_by(topic_id) |>
  summarise(total_words = sum(paragraph_word_count)) |>
  mutate(
    pct_total = total_words / sum(total_words) * 100
  )

p <- ggplot(df_topic_distribution, aes(x = topic_id, y = pct_total)) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = paste0(round(pct_total, 1), "%")),
    vjust = -0.5,
    size = 3
  ) +
  # coord_flip() +
  labs(
    x = "Topic",
    y = "Percentage of Total Words",
    title = "Distribution of Words across Topics"
  ) +
  theme_minimal()

ggsave(here("Outputs/06_plot_topic_distribution.png"))
write_rds(p, here("Outputs/06_plot_topic_distribution.rds"))
p
```

## Analysis of Topic Representation

Which businesses are most prevalent in each topic?

```{r}
#| label: topic-examples

# topics_bn <- 
  paragraphs_topics_wide |>
  # filter(topic != "-1") |>
  group_by(topic_id, Title) |>
  summarise(total_words = sum(paragraph_word_count), .groups = "drop") |>
  arrange(topic_id, desc(total_words)) |>
  group_by(topic_id) |>
  slice_head(n = 5) |>
  ungroup() |> 
  print(n = Inf)
```

Which climate keywords are most prevalent in each topic?

```{r}
# top_keywords_per_topic <- 
  paragraphs_topics_wide |>
  filter(
    keywords_found != ""#, 
    # topic != "-1"
  ) |>
  separate_rows(keywords_found, sep = ", ") |> 
  group_by(topic_id) |>
  count(keywords_found) |>
  mutate(
    freq = n / sum(n) * 100
  ) |>
  arrange(topic_id, desc(freq)) |> 
  slice_head(n = 5) |> 
  print(n = Inf)

# ggplot(
#   top_keywords_per_topic, 
#   aes(
#     x = reorder_within(keywords_found, freq, topic), 
#     y = freq, 
#     fill = topic)
#   ) +
#   geom_col(show.legend = FALSE) +
#   scale_x_reordered() +
#   coord_flip() +
#   facet_wrap(~ topic, scales = "free_y") +
#   labs(
#     title = "Top Keywords by Topic",
#     x = NULL,
#     y = "Count"
#   ) +
#   theme_minimal()
```

What are the most representative documents for each topic?


```{r}
documents_with_topics |>
  group_by(topic_id) |> 
  arrange(desc(topic_probability)) |>
  slice_sample(n = 5) |> 
  select(topic_id, transcript_text) |> 
  datatable()
```

## Party Distribution per Topic

Party composition across all topics (stacked bar charts) based on word counts.

```{r}
#| label: party-distribution-per-topic

df_plot <- paragraphs_topics_wide |>
  mutate(topic_id = factor(topic_id, levels = rev(sort(unique(topic_id))))) |> 
  # filter(topic != "-1") |>
  group_by(topic_id, party_abb) |>
  summarise(
    group_words = sum(paragraph_word_count),
    .groups = 'drop_last'
  ) |>
  mutate(
    pct_of_topic = group_words / sum(group_words) * 100
  ) |>
  drop_na()

p <- ggplot(
  df_plot,
  aes(x = pct_of_topic, y = topic_id, fill = party_abb)
) +
  geom_col(position = "stack") +
  geom_text(
    aes(
      label = ifelse(pct_of_topic > 15, sprintf("%.1f%%", pct_of_topic), "")
    ),
    position = position_stack(vjust = 0.5),
    color = "white",
    size = 3.5) +
  scale_fill_manual(values = party_colors) +
  labs(
    title = "Party Share per Topic",
    x = "Share (%)",
    y = "Topic",
    fill = "Party"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5)
  )

ggsave(here("Outputs/06_plot_topic_party_distribution.png"))
write_rds(p, here("Outputs/06_plot_topic_party_distribution.rds"))
p
```

# Heatmap of part-topic emphasis

x = share of party in parliament
y = share of party in topic
x/y = over/under emphasis of topic by party compared to overall parliamentary presence

```{r}
#| label: heatmap-party-topic

counts <- paragraphs_topics_wide |>
  # topic 6 and 7 can be joined into one topic, as they are very similar and both have low representation
  mutate(
    topic_id = case_when(
      topic_id %in% c(6,7) ~ "6",
      .default = as.character(topic_id)
    )
  ) |>
  group_by(party_abb, topic_id) |>
  summarise(
    wordcount_party_topic = sum(paragraph_word_count),
    .groups = "drop"
  ) |>
  left_join(
    paragraphs_topics_wide |>
      group_by(party_abb) |>
      summarise(wordcount_party = sum(paragraph_word_count), .groups = "drop"),
    by = "party_abb"
  )

shares <- counts |>
  group_by(topic_id) |>
  mutate(
    party_share_in_topic = wordcount_party_topic / sum(wordcount_party_topic) * 100
  ) |>
  ungroup() |>
  select(party_abb, topic_id, party_share_in_topic) |>
  pivot_wider(
    names_from = topic_id,
    names_prefix = "party_share_in_topic",
    values_from = party_share_in_topic
  ) |>
  left_join(
    paragraphs_topics_wide |>
      group_by(party_abb) |>
      summarise(
        wordcount_party = sum(paragraph_word_count),
        .groups = "drop"
      ) |>
      mutate(party_share_in_parl = wordcount_party / sum(wordcount_party) * 100) |>
      select(party_abb, party_share_in_parl),
    by = "party_abb"
  ) |>
  select(party_abb, party_share_in_parl, starts_with("party_share_in_topic"))

shares_long <- shares |>
  pivot_longer(
    cols = starts_with("party_share_in_topic"),
    names_to = "topic_id",
    values_to = "party_share_in_topic"
  ) |>
  mutate(
    topic_id = str_remove(topic_id, "party_share_in_topic"),
    diff_from_parl = party_share_in_topic - party_share_in_parl,
    topic_id = factor(as.numeric(topic_id), levels = rev(sort(unique(as.numeric(topic_id))))),
    party_abb = fct_rev(factor(party_abb))
  )

p <- ggplot(shares_long, aes(x = party_abb, y = topic_id, fill = diff_from_parl)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%+.1f", diff_from_parl)), size = 3) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = 0,
    name = "Deviation from\nparl. share (%)"
  ) +
  labs(
    x = "Party",
    y = "Topic",
    title = "Party share in topic vs. party share in parliament"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(here("Outputs/06_plot_heatmap_topic_party.png"))
write_rds(p, here("Outputs/06_plot_heatmap_topic_party.rds"))
p
```

# Regression Analysis

```{r}
#| label: regression-analysis-setup

# create dummy variables for topics 1 to 7
regression_data <- paragraphs_topics_wide |>
  distinct(transcript_id, topic_id, party_abb) |>
  mutate(
    topic_1 = factor(ifelse(topic_id == 1, 1, 0)),
    topic_2 = factor(ifelse(topic_id == 2, 1, 0)),
    topic_3 = factor(ifelse(topic_id == 3, 1, 0)),
    topic_4 = factor(ifelse(topic_id == 4, 1, 0)),
    topic_5 = factor(ifelse(topic_id == 5, 1, 0)),
    topic_6 = factor(ifelse(topic_id == 6, 1, 0)),
    topic_7 = factor(ifelse(topic_id == 7, 1, 0)),
    party_abb = relevel(factor(party_abb), ref = "FDP") # change according to hypothesis
  )
```

What influence does party affiliation have on topic assignment?

```{r}
#| label: regression-analysis

# logistic regression with topic 1 as dependent variable and party affiliation as independent variable
# model_1 <- glm(
#   topic_1 ~ party_abb,
#   data = regression_data,
#   family = binomial()
# )
# summary(model_1)

# for loop over topics 1 to 7
models <- list()
for (i in 1:7) {
  formula <- as.formula(paste0("topic_", i, " ~ party_abb"))
  model <- glm(formula, data = regression_data, family = binomial())
  models[[paste0("model_", i)]] <- model
}

# print summary of each model
for (i in 1:7) {
  cat(paste0("Model for Topic ", i, ":\n"))
  print(summary(models[[paste0("model_", i)]]))
  cat("\n\n")
}

```

# compare Legislative Periods

Create visualization like this:

boxplot combined with scatter points, where each point represents a party's share in a topic for a given legislative period.

on y axis: the diff_from_parl of all parties for a given topic
on x axis: legislative period

variables in data: topic_id, party_abb, legislative_period
variable to calculate: diff_from_parl

```{r}
counts <- paragraphs_topics_wide |> 
  group_by(party_abb, topic_id, legislative_period) |>
  summarise(
    wordcount_party_topic = sum(paragraph_word_count),
    .groups = "drop"
  ) |>
  left_join(
    paragraphs_topics_wide |>
      group_by(party_abb) |>
      summarise(wordcount_party = sum(paragraph_word_count), .groups = "drop"),
    by = "party_abb"
  )

shares <- counts |> 
  group_by(topic_id, legislative_period) |>
  mutate(
    party_share_in_topic = wordcount_party_topic / sum(wordcount_party_topic) * 100
  ) |>
  ungroup() |>
  select(party_abb, topic_id, legislative_period, party_share_in_topic) |>
  pivot_wider(
    names_from = topic_id,
    names_prefix = "party_share_in_topic",
    values_from = party_share_in_topic
  ) |>
  left_join(
    paragraphs_topics_wide |>
      group_by(party_abb, legislative_period) |>
      summarise(wordcount_party = sum(paragraph_word_count), .groups = "drop") |>
      group_by(legislative_period) |>
      mutate(party_share_in_parl = wordcount_party / sum(wordcount_party) * 100) |>
      select(party_abb, legislative_period, party_share_in_parl),
    by = c("party_abb", "legislative_period")
  ) |>
  select(party_abb, legislative_period, party_share_in_parl, starts_with("party_share_in_topic"))

shares_long <- shares |> 
  pivot_longer(
    cols = starts_with("party_share_in_topic"),
    names_to = "topic_id",
    values_to = "party_share_in_topic"
  ) |>
  mutate(
    topic_id = str_remove(topic_id, "party_share_in_topic"),
    diff_from_parl = party_share_in_topic - party_share_in_parl,
    topic_id = factor(as.numeric(topic_id), levels = rev(sort(unique(as.numeric(topic_id))))),
    party_abb = fct_rev(factor(party_abb))
  )

plots <- map(unique(shares_long$topic_id), function(topic) {
  df_plot <- shares_long |> 
    filter(topic_id == topic)
  
  ggplot(df_plot, aes(x = legislative_period, y = diff_from_parl)) +
    geom_boxplot() +
    geom_jitter(width = 0.2, height = 0, aes(shape = party_abb)) +
    labs(
      title = paste("Party Share Difference from Parliament by Topic", topic),
      x = "Legislative Period",
      y = "Difference from Parliament (%)",
      shape = "Party"
    ) +
    theme_minimal()
})

walk2(plots, unique(shares_long$topic_id), function(plot, topic) {
  ggsave(
    filename = here(paste0("Outputs/06_plot_period_comparison_topic_", topic, ".png")),
    plot = plot
  )
  write_rds(
    plot,
    here(paste0("Outputs/06_plot_period_comparison_topic_", topic, ".rds"))
  )
})
````