---
title: "Topic Analysis"
format: html
execute:
  warning: false
---

# Setup

Import packages and helper functions.

```{r}
library(tidyverse)
library(here)
library(knitr)
source(here("Scripts/00_functions.R"))
```

# Import Data

Read topic modeling results and join with paragraph-level data.

```{r}
documents_with_topics <- read_csv(here("Data/06_documents_with_topics.csv")) |>
  mutate(
    transcript_id = as.character(transcript_id),
    topic = sprintf("%02d", as.numeric(topic))
  ) |>
  filter(topic != "-01")

topic_info <- read_csv(here("Data/06_topic_info.csv")) |>
  rename(topic = Topic, topic_name = Name) |>
  mutate(topic = sprintf("%02d", as.numeric(topic))) |>
  filter(topic != "-01")

paragraphs_climate <- read_rds(here("Data/05_paragraphs_classified_wide_final.rds")) |>
  filter(final_climate == TRUE) |>
  mutate(
    ParlGroupName = case_when(
      ParlGroupAbbreviation %in% c("M-E", "BD") ~ "Die Mitte-Fraktion / CVP-& BD-Fraktion",
      .default = ParlGroupName
    )
  )
```

# Data Preparation

Join topic assignments with paragraph-level metadata.

```{r}
documents_with_topics_wide <- documents_with_topics |>
  left_join(topic_info, by = join_by(topic))

paragraphs_topics_wide <- documents_with_topics_wide |>
  inner_join(
    paragraphs_climate,
    by = join_by(transcript_id),
    relationship = "one-to-many"
  ) |>
  drop_na(ParlGroupAbbreviation)
```

# Visualization Setup

Define party display order and colour palette.

```{r}
party_names_order <- c(
  "Sozialdemokratische Fraktion",
  "Grüne Fraktion",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion",
  "Grünliberale Fraktion",
  "FDP-Liberale Fraktion",
  "Fraktion der Schweizerischen Volkspartei"
)

party_colors <- c(
  "Sozialdemokratische Fraktion" = "#E4032E",
  "Grüne Fraktion" = "#84B547",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion" = "#FF8C00",
  "Grünliberale Fraktion" = "#650959ff",
  "FDP-Liberale Fraktion" = "#0571C1",
  "Fraktion der Schweizerischen Volkspartei" = "#00843D"
)
```

# Topic Representation

Display topic labels generated by the representation models.

```{r}
topic_representation <- topic_info |>
  select(topic, OpenAI)

write_rds(topic_representation, here("Outputs/06_topic_representation.rds"))

topic_representation |>
  kable(
    caption = "Topic Representation based on OpenAI summaries",
    col.names = c("Topic", "Representation")
  )
```

# Topic Distribution

Overall distribution of paragraphs across topics.

```{r}
df_topic_distribution <- paragraphs_topics_wide |>
  group_by(topic) |>
  summarise(n_paragraphs = n()) |>
  mutate(
    topic = factor(topic, levels = rev(sort(unique(topic)))),
    pct_total = n_paragraphs / sum(n_paragraphs) * 100
  )

plot <- ggplot(df_topic_distribution, aes(x = topic, y = pct_total)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(
    label = paste0(round(pct_total, 1), "%")),
    hjust = -0.2, size = 3) +
  coord_flip() +
  labs(
    x = "Topic",
    y = "Percentage of Total Paragraphs",
    title = "Distribution of Paragraphs across Topics"
  ) +
  theme_minimal()

ggsave(here("Outputs/06_plot_topic_distribution.png"))
write_rds(plot, here("Outputs/06_plot_topic_distribution.rds"))
plot
```

# Party Distribution per Topic

Party composition across all topics (pie charts).

```{r}
df_plot <- paragraphs_topics_wide |>
  group_by(topic, ParlGroupName) |>
  summarise(
    group_count = n(),
    .groups = 'drop_last'
  ) |>
  mutate(
    pct_of_topic = group_count / sum(group_count) * 100,
    ParlGroupName = factor(ParlGroupName, levels = party_names_order)
  ) |>
  drop_na()

plot <- ggplot(
  df_plot,
  aes(x = "", y = pct_of_topic, fill = ParlGroupName)
) +
  geom_col() +
  coord_polar("y", start = 0) +
  facet_wrap(vars(topic)) +
  scale_fill_manual(values = party_colors) +
  labs(
    title = "Party Share per Topic",
    fill = ""
  ) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10, face = "bold")
  ) +
  guides(fill = guide_legend(nrow = 3))

ggsave(here("Outputs/06_plot_topic_party_distribution.png"))
write_rds(plot, here("Outputs/06_plot_topic_party_distribution.rds"))
plot
```

# Single Topic Visualization

Party composition for a specific topic.

```{r}
topic_to_visualize <- "01"

df_plot <- paragraphs_topics_wide |>
  filter(topic == topic_to_visualize) |>
  group_by(ParlGroupName) |>
  summarise(
    group_count = n(),
    .groups = 'drop_last'
  ) |>
  mutate(
    pct_of_topic = group_count / sum(group_count) * 100,
    ParlGroupName = factor(ParlGroupName, levels = party_names_order)
  ) |>
  drop_na()

plot <- ggplot(
  df_plot,
  aes(x = "", y = pct_of_topic, fill = ParlGroupName)
  ) +
  geom_col() +
  geom_text(aes(
    label = paste0(round(pct_of_topic, 1), "%")),
    position = position_stack(vjust = 0.5)
  ) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = party_colors) +
  labs(
    title = paste("Party Share for Topic", topic_to_visualize),
    fill = ""
  ) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    plot.title = element_text(hjust = 0.5)
  ) +
  guides(fill = guide_legend(nrow = 3))

ggsave(here(paste0("Outputs/06_plot_topic_", topic_to_visualize, ".png")))
write_rds(plot, here(paste0("Outputs/06_plot_topic_", topic_to_visualize, ".rds")))
plot
```
