---
title: "Speech Classification"
output-file: "Outputs/10_speech_classification.pdf"
format: pdf
execute:
  echo: false
  message: false
  warning: false
---

## Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
library(openxlsx)
library(tidytext)
library(stopwords)
library(here)
source(here("Scripts/00_functions.R"))
```

## Data Import

Read in all presaved data.

```{r}
#| label: data import

paragraphs_cleaned <- read_rds(here("Data/paragraphs_cleaned.rds"))
french_dictionary_curated <- read.xlsx(
  here("french_dictionary_curated.xlsx")) |>
  rename(keyword_original = keyword_french)
german_dictionary_curated <- read.xlsx(
  here("german_dictionary_curated.xlsx")) |>
  rename(keyword_original = keyword_german)
businesses_cleaned <- read_rds(
  here("Data/businesses_cleaned.rds"))
subjects <- read_rds(here("Data/subjects.rds"))
```

## Dictionary

Prepare the dictionary.

```{r}
#| label: dictionary

dictionary_all_cleaned <- german_dictionary_curated |>
  bind_rows(french_dictionary_curated) |>
  mutate(
    keyword_original = str_to_lower(keyword_original),
    keyword_original = str_replace_all(keyword_original, "[^[:alnum:]'-]", ""),
    keyword_original = str_squish(keyword_original)
  )

# save as rds
write_rds(
  dictionary_all_cleaned, 
  here("Data/dictionary_all_cleaned.rds"))
```

## Data Pre-Processing

Pre-process the data.

```{r}
#| label: data cleaning
paragraphs_words <- paragraphs_cleaned |>
  select(paragraph_id, paragraph) |>
  unnest_tokens(
    output = "word",
    input = "paragraph",
    token = "words",
    drop = FALSE
  )

# get german stopwords from stopwords package
# add additional stopwords
german_stopwords <- c(
  stopwords::stopwords("de", source = "snowball"),
  "dass"
) |>
  sort()

# get french stopwords from stopwords package
french_stopwords <- stopwords::stopwords("fr", source = "snowball")

# put together stopwords
stopwords_all <- tibble(stopwords = c(german_stopwords, french_stopwords))
# write_rds(stopwords_all, "Data/stopwords_all.rds")

paragraphs_words_cleaned <- paragraphs_words |>
  mutate(
    word = str_to_lower(word),
    word = str_replace_all(word, "[^[:alnum:]'-]", ""),
    word = str_squish(word)
  ) |>
  anti_join(stopwords_all, by = join_by(word == stopwords))

# save as rds
write_rds(paragraphs_words_cleaned, here("Data/paragraphs_words_cleaned.rds"))
```

## Dictionary Classification

Apply the Dictionary to the Data and count the number of matches.

```{r}
#| label: apply dictionary

paragraphs_classified <- paragraphs_words_cleaned |>
  left_join(dictionary_all_cleaned, by = join_by(word == keyword_original)) |>
  group_by(paragraph_id, paragraph) |>
  summarise(
    keywords_found = paste(na.omit(unique(keyword_english)), collapse = ", "),
    n_keywords_found = sum(!is.na(keyword_english)),
    climate_paragraph = if_else(n_keywords_found > 0, TRUE, FALSE),
    .groups = "drop"
  ) |>
  left_join(paragraphs_cleaned, by = join_by(paragraph_id))

# save as rds
write_rds(paragraphs_classified, here("Data/paragraphs_classified.rds"))
```

The classification happens in three steps. First define businesses that have climate paragraphs as climate_businesses.

```{r}
#| label: climate businesses

paragraphs_subjects <- left_join(
  paragraphs_classified,
  subjects,
  by = join_by(IdSubject),
  relationship = "many-to-many"
)

paragraphs_subjects_businesses <- left_join(
  paragraphs_subjects,
  businesses_cleaned,
  by = join_by(BusinessShortNumber),
  relationship = "many-to-one"
)

climate_businesses <- paragraphs_subjects_businesses |>
  filter(climate_paragraph == TRUE) |>
  distinct(BusinessShortNumber) |>
  pull(BusinessShortNumber)
```

Second define paragraphs that belong to a climate business as climate relevant.

```{r}
#| label: climate paragraphs

paragraphs_classified_wide <- paragraphs_subjects_businesses |>
  mutate(
    climate = case_when(
      BusinessShortNumber %in% climate_businesses ~ TRUE,
      .default = FALSE
    )
  ) |>
  rename(paragraph = paragraph.x)
```

Third define businesses that have at least 10% of the paragraphs that match the dictionary as climate_business_final.

```{r}
paragraphs_classified_wide_final <- paragraphs_classified_wide |>
group_by(BusinessShortNumber) |>
mutate(
  n_paragraphs = n(),
  n_climate_paragraphs = sum(climate_paragraph),
  proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
  final_climate = case_when(
    proportion_climate_paragraphs >= 0.1 ~ TRUE,
    .default = FALSE
  ),
  .groups = "drop"
)

write_rds(
  paragraphs_classified_wide_final, 
  here("Data/paragraphs_classified_wide_final.rds"))
```

# Visualize

Let's visualize the data. First show many paragraphs are climate relevant vs. non-relevant.

```{r}
plot_categorical_distribution(
  paragraphs_classified_wide_final,
  group_col = "final_climate",
  x_label = "Climate Relevant Paragraph",
  title = "Distribution of Climate Relevant Paragraphs",
  output_path = "Outputs/paragraphs_climate_distribution.png"
)
```

Show how many businesses are climate relevant vs. non-relevant

```{r}
df_grouped <- paragraphs_classified_wide_final |> 
  distinct(BusinessShortNumber, final_climate)

plot_categorical_distribution(
  df_grouped,
  group_col = "final_climate",
  x_label = "Climate Relevant Businesses",
  y_label = "Percentage of Businesses",
  label_suffix = "businesses",
  title = "Distribution of Climate Relevant Businesses",
  output_path = "Outputs/paragraphs_climate_distribution.png"
)
```

## R Code

For better readability, all the code from the document is added here:

```{r ref.label=knitr::all_labels()}
#| echo: true
#| eval: false

```