---
title: "Speech Classification"
format: pdf
execute:
  warning: false
---

## Setup

Import all necessary packages and functions.

```{r}
#| label: setup

library(tidyverse)
library(openxlsx)
library(tidytext)
library(stopwords)
library(here)
library(zoo)
source(here("Scripts/00_functions.R"))
```

## Data Import

Read in all presaved data.

```{r}
#| label: data import

paragraphs_cleaned <- read_rds(here("Data/paragraphs_cleaned.rds"))
french_dictionary_curated <- read.xlsx(
  here("french_dictionary_curated.xlsx")) |>
  rename(keyword_original = keyword_french)
german_dictionary_curated <- read.xlsx(
  here("german_dictionary_curated.xlsx")) |>
  rename(keyword_original = keyword_german)
businesses_cleaned <- read_rds(
  here("Data/businesses_cleaned.rds"))
subjects <- read_rds(here("Data/subjects.rds"))
```

## Dictionary

Prepare the dictionary.

```{r}
#| label: dictionary

dictionary_all_cleaned <- german_dictionary_curated |>
  bind_rows(french_dictionary_curated) |>
  mutate(
    keyword_original = str_to_lower(keyword_original),
    keyword_original = str_replace_all(keyword_original, "[^[:alnum:]'-]", ""),
    keyword_original = str_squish(keyword_original)
  )

# save as rds
write_rds(
  dictionary_all_cleaned, 
  here("Data/dictionary_all_cleaned.rds"))
```

## Data Pre-Processing

Pre-process the data.

```{r}
#| label: data cleaning
paragraphs_words <- paragraphs_cleaned |>
  select(paragraph_id, paragraph) |>
  unnest_tokens(
    output = "word",
    input = "paragraph",
    token = "words",
    drop = FALSE
  )

# get german stopwords from stopwords package
# add additional stopwords
german_stopwords <- c(
  stopwords::stopwords("de", source = "snowball"),
  "dass"
) |>
  sort()

# get french stopwords from stopwords package
french_stopwords <- stopwords::stopwords("fr", source = "snowball")

# put together stopwords
stopwords_all <- tibble(stopwords = c(german_stopwords, french_stopwords))
# write_rds(stopwords_all, "Data/stopwords_all.rds")

paragraphs_words_cleaned <- paragraphs_words |>
  mutate(
    word = str_to_lower(word),
    word = str_replace_all(word, "[^[:alnum:]'-]", ""),
    word = str_squish(word)
  ) |>
  anti_join(stopwords_all, by = join_by(word == stopwords))

# save as rds
write_rds(paragraphs_words_cleaned, here("Data/paragraphs_words_cleaned.rds"))
```

## Dictionary Classification

Apply the Dictionary to the Data and count the number of matches.

```{r}
#| label: apply dictionary

paragraphs_classified <- paragraphs_words_cleaned |>
  left_join(dictionary_all_cleaned, by = join_by(word == keyword_original)) |>
  group_by(paragraph_id, paragraph) |>
  summarise(
    keywords_found = paste(na.omit(unique(keyword_english)), collapse = ", "),
    n_keywords_found = sum(!is.na(keyword_english)),
    climate_paragraph = if_else(n_keywords_found > 0, TRUE, FALSE),
    .groups = "drop"
  ) |>
  left_join(paragraphs_cleaned, by = join_by(paragraph_id))

# save as rds
write_rds(paragraphs_classified, here("Data/paragraphs_classified.rds"))
```

The classification happens in three steps. First define businesses that have climate paragraphs as climate_businesses.

```{r}
#| label: climate businesses

paragraphs_subjects <- left_join(
  paragraphs_classified,
  subjects,
  by = join_by(IdSubject),
  relationship = "many-to-many"
)

paragraphs_subjects_businesses <- left_join(
  paragraphs_subjects,
  businesses_cleaned,
  by = join_by(BusinessShortNumber),
  relationship = "many-to-one"
)

climate_businesses <- paragraphs_subjects_businesses |>
  filter(climate_paragraph == TRUE) |>
  distinct(BusinessShortNumber) |>
  pull(BusinessShortNumber)
```

Second define paragraphs that belong to a climate business as climate relevant.

```{r}
#| label: climate paragraphs

paragraphs_classified_wide <- paragraphs_subjects_businesses |>
  mutate(
    climate = case_when(
      BusinessShortNumber %in% climate_businesses ~ TRUE,
      .default = FALSE
    )
  ) |>
  rename(paragraph = paragraph.x)
```

Third define businesses that have at least 10% of the paragraphs that match the dictionary as final_climate.

```{r}
paragraphs_classified_wide_final <- paragraphs_classified_wide |>
group_by(BusinessShortNumber) |>
mutate(
  n_paragraphs = n(),
  n_climate_paragraphs = sum(climate_paragraph),
  proportion_climate_paragraphs = n_climate_paragraphs / n_paragraphs,
  final_climate = case_when(
    proportion_climate_paragraphs >= 0.1 ~ TRUE,
    .default = FALSE
  ),
  .groups = "drop"
)

write_rds(
  paragraphs_classified_wide_final, 
  here("Data/paragraphs_classified_wide_final.rds"))
```

Important: The climate related debates are therefore based on the nature of the discussed parliamentary business. If the business is considered climate related, all debates around this business are considered climate related.

# Visualize

Let's visualize the data. First show many paragraphs are climate relevant vs. non-relevant.

```{r}
p <- plot_categorical_distribution(
  paragraphs_classified_wide_final,
  group_col = "final_climate",
  x_label = "Climate Relevant Paragraph",
  title = "Distribution of Climate Relevant Paragraphs",
  output_path = "Outputs/10_paragraphs_climate_distribution.png"
)
write_rds(p, here("Outputs/10_paragraphs_climate_distribution.rds"))

# show the plot
p
```

Show how many businesses are climate relevant vs. non-relevant

```{r}
df_grouped <- paragraphs_classified_wide_final |> 
  distinct(BusinessShortNumber, final_climate)

p <- plot_categorical_distribution(
  df_grouped,
  group_col = "final_climate",
  x_label = "Climate Relevant Businesses",
  y_label = "Percentage of Businesses",
  label_suffix = "businesses",
  title = "Distribution of Climate Relevant Businesses",
  output_path = "Outputs/10_paragraphs_climate_distribution.png"
)
write_rds(p, here("Outputs/10_paragraphs_climate_distribution.rds"))

# show the plot
p
```

In absolute terms, the number of climate relevant paragraphs in comparison to non-climate paragraphs developed like this.

```{r}
# in absolute terms
df_grouped <- paragraphs_classified_wide_final |> 
  mutate(meeting_date = as.Date(Start)) |> 
  group_by(meeting_date, final_climate) |> 
  summarise(count = n(), .groups = "drop")

p <- ggplot(df_grouped, aes(x = meeting_date, y = count, color = final_climate)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Climate vs. Non-Climate Paragraphs over Time",
    x = "Date",
    y = "Number of Paragraphs",
    color = "Climate"
  ) +
  theme_minimal()
ggsave(here("Outputs/10_plot_timeseries_absolute.png"))
write_rds(p, here("Outputs/10_plot_timeseries_absolute.rds"))

# show the plot
p
```

Show how the proportion of climate relevant paragraphs developed over the last 10 years.

```{r}
# in relative terms
share <- paragraphs_classified_wide_final |> 
  mutate(meeting_date = as.Date(Start)) |> 
  group_by(meeting_date) |> 
  summarise(share_climate = mean(final_climate) * 100) |> 
  arrange(meeting_date) |> 
  mutate(rolling_avg = rollmean(share_climate, k = 30, fill = NA))

p <- ggplot(share, aes(x = meeting_date)) +
  geom_line(aes(y = share_climate), alpha = 0.3, color = "gray50") +
  geom_line(aes(y = rolling_avg), color = "steelblue", size = 1) +
    labs(
    title = "Share of Climate Related Debates",
    x = "Date",
    y = "Share (%)"
  ) +
  theme_minimal()
ggsave(here("Outputs/10_plot_timeseries_relative.png"))
write_rds(p, here("Outputs/10_plot_timeseries_relative.rds"))

# show the plot
p
```

Let's see which parties have been most active in climate debates over time.

```{r}
party_names_order <- c(
  "Sozialdemokratische Fraktion",
  "Grüne Fraktion", 
  "Die Mitte-Fraktion / CVP-Fraktion",
  "Grünliberale Fraktion",
  "FDP-Liberale Fraktion",
  "Fraktion der Schweizerischen Volkspartei"
)
party_colors <- c(
  "Sozialdemokratische Fraktion" = "#E4032E",
  "Grüne Fraktion" = "#84B547",
  "Die Mitte-Fraktion / CVP-& BD-Fraktion" = "#FF8C00",
  "Grünliberale Fraktion" = "#650959ff",
  "FDP-Liberale Fraktion" = "#0571C1",
  "Fraktion der Schweizerischen Volkspartei" = "#00843D"
)

share <- paragraphs_classified_wide_final |> 
  filter(final_climate == TRUE) |> 
  mutate(
    meeting_date = as.Date(Start),
    ParlGroupName = case_when(
      ParlGroupAbbreviation == "M-E|BD" ~ "Die Mitte-Fraktion / CVP-& BD-Fraktion",
      .default = ParlGroupName
    )
  ) |> 
  group_by(meeting_date) |> 
  mutate(total_climate = n()) |> 
  group_by(meeting_date, ParlGroupName) |> 
  summarise(
    share_party = (n() / first(total_climate)) * 100,
    .groups = "drop") |> 
  mutate(
    ParlGroupName = factor(ParlGroupName, levels = party_names_order)) |>
  arrange(meeting_date, ParlGroupName) |> 
  group_by(ParlGroupName) |> 
  mutate(rolling_avg = rollmean(share_party, k = 30, fill = NA)) |> 
  drop_na(ParlGroupName)

p <- ggplot(share, aes(x = meeting_date, y = rolling_avg, color = ParlGroupName)) +
  geom_line(size = 1) +
  scale_color_manual(values = party_colors) +
  labs(
    title = "Share of Climate Related Debates per Party",
    x = "Date",
    y = "Share (%)",
    color = "Party Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9)
  ) +
  guides(color = guide_legend(nrow = 2))
ggsave(here("Outputs/10_plot_timeseries_per_party.png"), width = 10, height = 8)
write_rds(p, here("Outputs/10_plot_timeseries_per_party.rds"))

# show the plot
p
```

There was a time when the mainstream right and populist right parties dominated the climate debate until around 2021, after which it became much more balanced between the parties, with the Greens having the largest share of the discussion.

Interestingly, the proportion of speeches made by the Greens and Green Liberals diverges sharply from 2020 onwards.

The graph does not take into account the size of the parliamentary groups. On average, the mainstream left and mainstream right have a larger share than the Greens and Green Liberals, with the right-wing populist SVP having the largest share in parliament (Source: https://www.parlament.ch/de/organe/fraktionen/).

